---
title: "Does endophyte composition explain decay?"
author: "Marissa Lee"
date: "12/2/2018"
output: github_document
---

Contents:
1. Summarize sequencing effort
2. Create a filtered dataset without rare taxa
3. Identify taxa by functional groups
4. Export OTU tables for analysis
5. Summarize alpha diversity

___________________
___________________

Load libraries and functions
```{r}
#chunk options
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)

#---------------------------#
#libraries
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr
library(tidymodels) # broom
library(sjstats) #provides function eta_sq() to extract r2 for each predictor in a lm
library(grid) #plotting
library(gridExtra) # plotting
#library(GGally) # bivariate plots
#library(lmtest) # for coxtest()


#---------------------------#
#fxns
source('code/helper_fxns.R')
sourceDir('code')

```

Load decay data
```{r loaddata}

# samples and decay
stemSamples <- load_stemSamples() # this function is in load_microbeData.R
initial_mass <- read_in_initial_mass()
harvest_mass <- LoadHarvestFiles()
pmr <- Calc_massRemaining(initial_mass, harvest_mass) # calculate percent mass remaining (pmr)
pmr_byStem <- AvgPMR_byStem(pmr, long.form = F) # average pmr for each stem and timepoint
stem.respVars <- list("time7", "time13", "time25", "time37","time59")
decayfits <- read_csv("derived_data/decayfits.csv") # see 1_decayPatterns.Rmd to update
code.respVars <- list("w.t50","beta","alpha","w.r2")
# weibull params
# beta = scale param = higher values mean *slower* decay
# alpha = shape param = higher values mean more S-shaped
```

Load wood trait data
```{r}
traits.code <- read.csv(file = "derived_data/traits_code.csv", row.names = 1)
traits.stem <- read.csv(file = "derived_data/traits_stem.csv", row.names = 1)

# ordering and stuff
trait.order <- traitcol.order()
traitVars1 <- c(trait.order$phys.cols, trait.order$nutr.cols)
traitVars1 # slice1 (w/ size)
traitVars2 <- c(traitVars1, trait.order$cfract.cols)
traitVars2 # slice2 (w/o size)

```

Load residuals from 2b_woodTraits_explainDecay.Rmd
```{r}
resids_explainDecay.list <- readRDS(file = "derived_data/resids_explainDecay_list.RData")

```

Load microbial community data
```{r}
# samples and decay
stemSamples <- load_stemSamples() # this function is in load_microbeData.R

# microbes
microb.data <- load_MicrobeCollection(stemSamples)
taxAndFunguild <- microb.data$taxAndFunguild
comm.otu <- microb.data$comm.otu
seqSamples <- microb.data$seqSamples

#check for oomycetes
colnames(comm.otu)[grep("oo", colnames(comm.otu))] # 9
taxAndFunguild %>%
  filter(kingdom != "Fungi") #9

```

___________________
# 1. Summarize sequencing effort

Sample-effort curves
```{r}
#this takes a while... uncomment if the data changes
#plot_sampleEffortCurves(comm.otu) # output/figures/sampleEffortCurve.pdf

```

How many total OTUs and reads?
```{r}
# how many OTUs?
dim(comm.otu)
dim(taxAndFunguild)
taxAndFunguild %>%
  group_by(kingdom) %>%
  summarize(n = length(OTUId))

# how many reads?
x <- grepl("oo", colnames(comm.otu))
comm.otu.oo <- comm.otu[,x]
comm.otu.fun <- comm.otu[,!x]

sum(colSums(comm.otu.oo))
sum(colSums(comm.otu.fun))

```

How many per sample?
```{r}
# mean number of reads per sample
mean(rowSums(comm.otu))
sd(rowSums(comm.otu))/sqrt(length(rowSums(comm.otu)))

# mean numbr of OTUs per sample
comm.otu.pa <- (comm.otu > 0) * 1
mean(rowSums(comm.otu.pa))
sd(rowSums(comm.otu.pa))/sqrt(length(rowSums(comm.otu.pa)))
```

___________________
# 2. Create a filtered dataset without rare taxa

Filter community matrix to include only taxa that are present in a least 20% of all the samples. This step removes taxa that may not contribute much to our understanding of the relationship between species’ multivariate abundance and environment.
```{r}
comm.otu.trimmed <- removeRareOTUs(comm.otu)

# are oomycetes in here? nope
colnames(comm.otu.trimmed)[grep("oo", colnames(comm.otu.trimmed))]

```
This drops all the oomycetes

___________________
# 3. Identify taxa by functional groups

How many OTUs matched to a genus-level taxonomic assignment?
```{r}
colnames(taxAndFunguild)
n.unc <- sum(taxAndFunguild$genus == "unclassified")
n.unc / dim(taxAndFunguild)[1]
taxAndFunguild %>%
  filter(genus != "unclassified") %>%
  group_by(Confidence.Ranking) %>%
  summarize(n = length(OTUId)) -> tmp
conf <- 244 + 54
conf/sum(tmp$n)
```

Identify the saprotrophs
```{r}
taxAndFunguild %>%
  filter(grepl("Sapro", Guild)) -> sapro.tax
  
comm.otu.sapro <- comm.otu[, colnames(comm.otu) %in% sapro.tax$OTUId]
dim(comm.otu.sapro) # 176 OTUs

comm.otu.sapro.trimmed <- comm.otu.trimmed[, colnames(comm.otu.trimmed) %in% sapro.tax$OTUId]
dim(comm.otu.sapro.trimmed) # only 11 OTUs

```
There are 176 saprotroph OTUs, of those 11 are present in the trimmed community

Identify the basidios
```{r}
taxAndFunguild %>%
  filter(grepl("Basidio", phylum)) -> basid.tax
basid.tax  
comm.otu.basid <- comm.otu[, colnames(comm.otu) %in% basid.tax$OTUId]
dim(comm.otu.basid) # 526 OTUs

comm.otu.basid.trimmed <- comm.otu.trimmed[, colnames(comm.otu.trimmed) %in% basid.tax$OTUId]
dim(comm.otu.basid.trimmed) # only 15 OTUs

```
There are 526 Basidiomycota OTUs, of those 15 are present in the trimmed community

Identify the oomycetes
```{r}
taxAndFunguild %>%
  filter(grepl("Protist", kingdom)) -> oo.tax
oo.tax  
comm.otu.oo <- comm.otu[, colnames(comm.otu) %in% oo.tax$OTUId]
dim(comm.otu.oo) # 9 OTUs
#remove empty samples
#comm.otu.oo <- comm.otu.oo[rowSums(comm.otu.oo) > 0,]

comm.otu.oo.trimmed <- comm.otu.trimmed[, colnames(comm.otu.trimmed) %in% oo.tax$OTUId]
dim(comm.otu.oo.trimmed) # 0 OTUs

```
There are 9 Oomycete OTUs, of those 0 are present in the trimmed community

Identify the Pathotrophs
```{r}
taxAndFunguild %>%
  filter(grepl("Patho", Guild)) -> patho.tax
  
comm.otu.patho <- comm.otu[, colnames(comm.otu) %in% patho.tax$OTUId]
dim(comm.otu.patho) # 129 OTUs

comm.otu.patho.trimmed <- comm.otu.trimmed[, colnames(comm.otu.trimmed) %in% patho.tax$OTUId]
dim(comm.otu.patho.trimmed) # only 11 OTUs

```
There are 129 Pathotroph OTUs, of those 10 are present in the trimmed community

___________________
# 4. Export OTU tables for analysis

```{r}

#all
dim(comm.otu) # 3297 OTUs
dim(comm.otu.trimmed) # 146 OTUs

#sapros
dim(comm.otu.sapro) # 176 OTUs
dim(comm.otu.sapro.trimmed) # only 11 OTUs

#basidios
dim(comm.otu.basid) # 526 OTUs
dim(comm.otu.basid.trimmed) # only 15 OTUs

#oomycetes
dim(comm.otu.oo) # 9 OTUs in 11 samples
dim(comm.otu.oo.trimmed) # 0 OTUs

#pathos
dim(comm.otu.patho) # 129 OTUs
dim(comm.otu.patho.trimmed) # only 10 OTUs


otu.list <- list(all = comm.otu,
                 sapro = comm.otu.sapro,
                 basid = comm.otu.basid,
                 oo = comm.otu.oo,
                 patho = comm.otu.patho)
saveRDS(otu.list, file = "derived_data/otu_list.RData")

otu.list.trimmed <- list(all = comm.otu.trimmed,
                 sapro = comm.otu.sapro.trimmed,
                 basid = comm.otu.basid.trimmed,
                 patho = comm.otu.patho.trimmed)
saveRDS(otu.list.trimmed, file = "derived_data/otu_list_trimmed.RData")
# remember there are 0 oomycete OTUs in the trimmed dataset so these are not included in the list


```

___________________
# 5. Summarize alpha diversity




