---
title: "Carbon fraction analyses"
author: "Marissa Lee"
date: "7/19/2019"
output: pdf_document
---

```{r, include = F}
#chunk options
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

```{r load}
# load libraries and fxns

#libraries
library(tidyverse) #data formatting
library(grid) #plotting
library(gridExtra) #plotting
library(corrplot) #plotting
#devtools::install_github("cornwell-lab-unsw/litterfitter") #fit decay models to timeseries
library(litterfitter) #fit decay models to timeseries
#library(lmtest) # for coxtest()
library(vegan) #diversity metrics
#library(rioja) #WA-PLS
#library(GGally)

#fxns
source('code/helper_fxns.R')
sourceDir('code')

```

Load C fraction data
```{r}

#load_Cfract<-function(){
  
  cfract <- read_csv(file = "data/Marissa_Chemistry_Data_export.csv")
  cfract <- cfract[complete.cases(cfract),] # get rid of empty row from the excel file
  
  #I think there is a typo and JASC2 was entered twice instead of JASC3. My original datasheet that I sent included JASC3. 
  #It would be good to double check this with Shawn's lab.
  
  # fix typo
  cfract %>%
    mutate(codeStem = ifelse(codeStem == "JASC2" &  labID %in% c(47,48), 
                             "JASC3", 
                             codeStem)) -> cfract
  
#plot
num.out <- length(unique(cfract$labID))/2
cfract %>%
    mutate(rep = paste0("rep",rep(c(1,2), times = num.out))) %>%
    select(codeStem, rep, perc.Total.lignin, Arabinose, Rhamnose, Galactose, Glucose, Xylose, Mannose, Total) %>%
    gather(key = "type", value = "value", -c(codeStem, rep)) %>%
    separate(codeStem, into = c("code","Stem"), sep = 4, remove = F) %>%
    mutate(size = ifelse(code == tolower(code), "small","large")) %>%
    rename('unique'=codeStem) -> tmp

# order code by percTotal.lignin
tmp %>%
  filter(type == "perc.Total.lignin") %>%
  group_by(code) %>%
  summarize(mean = mean(value)) %>%
  arrange(mean) -> code.order
tmp$code <- factor(tmp$code, levels = code.order$code)

p <- ggplot(tmp, aes(x = code, y = value, color = Stem)) +
  geom_point() +
  facet_wrap(~type, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
p
ggsave(p, filename = "cfract_distrb.pdf")

```

Check for analytical reps that are way different
```{r}
  
  #Look for analytical reps that are way different
cfract %>%
    mutate(rep = paste0("rep",rep(c(1,2), times = num.out))) %>%
    select(codeStem, rep, perc.Total.lignin, Arabinose, Rhamnose, Galactose, Glucose, Xylose, Mannose, Total) %>%
    gather(key = "type", value = "value", -c(codeStem, rep)) %>%
    spread(key = rep, value = value) %>%
    mutate(diff = abs(rep1 - rep2)) -> tmp
  
  #plot difference between analytical rep1 and rep2
  ggplot(tmp, aes(x = codeStem, y = diff)) +
    geom_point() +
    facet_wrap(~type, scales = "free_y") +
    theme_bw() +
    ylab("Abs. diff between two analytical reps") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```  
Go back and investigate the JASC/Arabinose and ACPA/Xylose outliers. 

Average the analytical reps
```{r}
  # average the 2 analytical reps
num.out <- length(unique(cfract$labID))/2
  cfract %>%
    mutate(rep = paste0("rep",rep(c(1,2), times = num.out))) %>%
    select(codeStem, rep, perc.Total.lignin, Arabinose, Rhamnose, Galactose, Glucose, Xylose, Mannose, Total) %>%
    gather(key = "type", value = "value", -c(codeStem, rep)) %>%
    spread(key = rep, value = value) %>%
    mutate(diff = abs(rep1 - rep2)) %>%
    mutate(avg = (rep1 + rep2) / 2) %>%
    select(codeStem, type, avg) %>%
    spread(key = type, value = avg) -> cfract.w
  
  # add unique, code, size, Stem
  cfract.w %>%
    separate(codeStem, into = c("code","Stem"), sep = 4, remove = F) %>%
    mutate(size = ifelse(code == tolower(code), "small","large")) %>%
    rename('unique'=codeStem) -> cfract.wt
  dim(cfract.wt)
  
#  return(cfract.wt)
#}
```

Add existing trait data to this C fraction data. Exclude two samples for which there isn't waterperc data. Remember that there is no stem-level barkthick or density data for large stem samples.
```{r}
# load all trait data
trait.data.l <- mergeTraitData()
trait.data.l %>%
  mutate(codeStem = paste0(code,Stem)) -> trait.data.l

# subset by the codeStem ids with Cfract data
select.codeStems <- unique(cfract.wt$unique)
trait.data.l %>%
  filter(codeStem %in% select.codeStems) -> select.trait.data.l

# make the dataframe wide so that each trait is a column
select.trait.data.l %>%
  spread(key = trait, value = trait.val) -> tmp
# need to create an average waterperc for each codeStem so that this data is aggregated at the same scale as the XRF trait data
tmp %>%
  group_by(codeStem) %>%
  summarize(mean.waterperc = mean(waterperc, na.rm =T)) -> mean.waterperc
tmp %>%
  select(-waterperc) %>%
  filter(!is.na(C)) %>%
  left_join(mean.waterperc) -> select.trait.data.w
dim(select.trait.data.w)

# remove waterperc with NAs (2 codeStems)
select.trait.data.w %>%
  filter(!is.nan(mean.waterperc)) %>%
  rename('waterperc' = 'mean.waterperc') -> select.trait.data.w
dim(select.trait.data.w)

# merge the C fract data with the other trait data
select.trait.data.w %>%
  left_join(cfract.wt) %>%
  select(-c(Total, unique, compositeSample)) -> trait.data.w
```

Check out correlations among C fractions and other traits. 
```{r}
#grab just the matrix of traits
colnames(trait.data.w)
trait.data.w %>%
  select(-c(code, size, Stem, codeStem)) -> matonly
matonly.s <- scale(as.matrix(matonly))

# plot correlations
pdf(file = "output/corplot.pdf")
corrplot(cor(matonly.s), type = "lower")
dev.off()

cor(matonly.s)

```

Kick out traits that are highly correlated
- C and perc.Total.lignin are highly positive correlated (0.8); keep perc.Total.lignin
- Glucose+Xylose and perc.Total.lignin are highly negatively correlated (-.8, -.7); keep perc.Total.lignin.
```{r}
#update
trait.data.w %>%
  select(-c(Glucose, Xylose, C)) -> trait.data.w.sel
trait.data.w.sel %>%
  select(-c(code, size, Stem, codeStem)) -> matonly
matonly.s <- scale(as.matrix(matonly))

# plot correlations again
corrplot(cor(matonly.s), type = "lower")
cor(matonly.s)

```
Note that waterperc and Mannose are negatively correlated (-0.62), but I'm leaving them in there.

Visualize the codeStems in trait space
```{r}
pca <- prcomp(matonly.s)
summary(pca)

# add back the sample info
df <- data.frame(pca$x[,1:2], trait.data.w.sel[,c("codeStem","code","size","Stem")])

#plot
datapc <- data.frame(varnames=rownames(pca$rotation), pca$rotation[,1:2])
mult <- min(
  (max(df[,"PC1"]) - min(df[,"PC2"])/(max(datapc[,"PC2"])-min(datapc[,"PC2"]))),
  (max(df[,"PC2"]) - min(df[,"PC1"])/(max(datapc[,"PC1"])-min(datapc[,"PC1"])))
)
datapc <- transform(datapc,
                    v1 = .7 * mult * datapc$PC1,
                    v2 = .7 * mult * datapc$PC2)
p <- ggplot(data=datapc, aes(x=v1, y=v2, label=varnames)) + 
  geom_text(size=3, vjust=1, color=1) + 
  geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),
               arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color=1) + 
  coord_fixed() +
  geom_text(mapping=aes(x=PC1, y=PC2, 
                        color = code, label = codeStem), data=df, inherit.aes = FALSE) +
  theme_bw() + xlab("PC1 (24%)") + ylab("PC2 (16%)") +
  guides(color = F)
p
ggsave(p, filename = "output/stem_traitord.pdf")

```

How do wood C fractions contribute to explaining decay?

(1) by codeStem

Merge trait matrix with the percent mass loss data
```{r}

# stem mass
initial_mass <- read_in_initial_mass()
harvest_mass <- LoadHarvestFiles()
pmr <- Calc_massRemaining(initial_mass, harvest_mass) # calculate percent mass remaining (pmr)
pmr_byStem <- AvgPMR_byStem(pmr) # average pmr for each stem and timepoint

#merge with trait data
trait.data.w %>%
  left_join(pmr_byStem) -> pmr.trait.data

```
Remember there are 34 codeStems/observations.

Set up models
```{r}
orig.x.vars <- c("Ca","Fe","K","Mn","N","P","Zn","waterperc")

cfract.x.vars <- c("Arabinose","Galactose","Mannose","perc.Total.lignin","Rhamnose","Glucose", "Xylose")
stem.respVars <- list("time7", "time13", "time25", "time37","time59")

pmr.trait.data %>%
  select(as.character(stem.respVars)) -> y.mat

rhs.orig <- paste(orig.x.vars, collapse = " + ")
rhs.plusC <- paste(c(orig.x.vars, cfract.x.vars), collapse = " + ")
rhs.Conly <- paste(c(cfract.x.vars), collapse = " + ")
```

(a) Original traits: "Ca","Fe","K","Mn","N","P","Zn","waterperc" predict percent mass remaining (PMR) at each time point
```{r}
# rhs.orig model
rhs.orig <- paste(orig.x.vars, collapse = " + ")
rhs.mat <- pmr.trait.data[,colnames(pmr.trait.data) %in% orig.x.vars]
rhs.mat.s <- scale(rhs.mat)

ModelFit_fxn <- function(y, y.data, rhs, rhs.data){
  
  #merge data
  data <- cbind(y.data, rhs.data)
  
  #create model formula
  string <- paste(y, " ~ ", rhs)
  fmla<-as.formula(string)
  mod.full<-lm(formula=fmla, data=data)
  
  return(mod.full)
}

# set up all the models
mod.full.list <- lapply(stem.respVars, function(x) {
  result <- ModelFit_fxn(y = x, y.data = y.mat,
                         rhs = rhs.orig, rhs.data = rhs.mat.s)
  return(result)
})

#do stepwise model selection
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="backward")
  return(mod.select)
  })

#pull coefs and r2s
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(sum.list, unlist(stem.respVars))
fitstat.df<-PullLmFitStats(sum.list, unlist(stem.respVars))
fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
coefs.df %>%
  left_join(r2.df) -> coefs.df

# which traits were included in the best models?
coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p.stem <- ggplot(tmp, aes(x = respvar, y = term, fill = r.squared)) +
  geom_tile(color = "black")
p.stem

# what direction are the associations?
p.stem <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.stem

```

Visualize codeStems in trait space, add pmr vectors
```{r}

#create ordination
pca <- prcomp(rhs.mat.s)
summary(pca)

#do envfit
pca.envfit <- envfit(pca, y.mat, na.rm = T)

# make plot
 # collect site scores from mod.obj
scrs <- as.data.frame(scores(pca, display = "sites"))
scrs <- cbind(scrs, codeStem = pmr.trait.data$codeStem)
scrs %>%
  left_join(pmr.trait.data) -> scrs
  
# collect vector scores from envfit.obj
spp.scrs <- as.data.frame(scores(pca.envfit, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), pval = pca.envfit$vectors$pvals)
  
# collect trait loadings
datapc <- data.frame(varnames=rownames(pca$rotation), pca$rotation[,1:2])
mult <- min(
  (max(df[,"PC1"]) - min(df[,"PC2"])/(max(datapc[,"PC2"])-min(datapc[,"PC2"]))),
  (max(df[,"PC2"]) - min(df[,"PC1"])/(max(datapc[,"PC1"])-min(datapc[,"PC1"])))
)
datapc <- transform(datapc,
                    v1 = .7 * mult * datapc$PC1,
                    v2 = .7 * mult * datapc$PC2)
# plot
mult <- 3 #multiplier for the arrows and text for envfit
mult.text <- 3
p <- ggplot(scrs) +
  
  # traits loadings
  geom_text(data = datapc, aes(x=v1, y=v2, label = varnames), 
            size=3, vjust=1, color=1) + 
  geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),
               arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color=1) +
  
  # sites
  geom_point(data = scrs, mapping = aes(x = PC1, y = PC2, color = code), alpha = .5) +
  coord_fixed() + ## need aspect ratio of 1!
  
  # envfit
  geom_segment(data = spp.scrs,
               aes(x = 0, xend = mult*PC1, y = 0, yend = mult*PC2),
               arrow = arrow(length = unit(0.2, "cm")), color = 'red') +
  geom_text(data = spp.scrs, 
            aes(x = mult.text*PC1, y = mult.text*PC2, 
                label = Species), 
            size = 4, fontface = "bold", color = "red") +
  xlab("PCoA 1") + ylab("PCoA 2")
p
ggsave(p, filename = "output/stem_pmr_traitord.pdf")

```

(b) Original traits plus C fractions...
```{r}
# rhs.orig model
rhs.plusC <- paste(c(orig.x.vars, cfract.x.vars), collapse = " + ")
rhs.mat <- pmr.trait.data[,colnames(pmr.trait.data) %in% c(orig.x.vars, cfract.x.vars)]
rhs.mat.s <- scale(rhs.mat)

# set up all the models
stem.respVars
mod.full.list <- lapply(stem.respVars, function(x) {
  result <- ModelFit_fxn(y = x, y.data = y.mat,
                         rhs = rhs.plusC, rhs.data = rhs.mat.s)
  return(result)
})

#do stepwise model selection
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="backward")
  return(mod.select)
  })

#pull coefs and r2s
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(mod.select.list, unlist(stem.respVars))
fitstat.df<-PullLmFitStats(mod.select.list, unlist(stem.respVars))

fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
r2.df
coefs.df %>%
  left_join(r2.df) -> coefs.df

# which traits were included in the best models?
coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p.stem <- ggplot(tmp, aes(x = respvar, y = term, fill = r.squared)) +
  geom_tile(color = "black")
p.stem

# what direction are the associations?
p.stem <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.stem


p.stem <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.stem

```

Visualize codeStems in trait space, add pmr vectors
```{r}

#create ordination
pca <- prcomp(rhs.mat.s)
summary(pca)

#do envfit
pca.envfit <- envfit(pca, y.mat, na.rm = T)

# make plot
 # collect site scores from mod.obj
scrs <- as.data.frame(scores(pca, display = "sites"))
scrs <- cbind(scrs, codeStem = pmr.trait.data$codeStem)
scrs %>%
  left_join(pmr.trait.data) -> scrs
  
# collect vector scores from envfit.obj
spp.scrs <- as.data.frame(scores(pca.envfit, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), pval = pca.envfit$vectors$pvals)
  
# collect trait loadings
datapc <- data.frame(varnames=rownames(pca$rotation), pca$rotation[,1:2])
mult <- min(
  (max(df[,"PC1"]) - min(df[,"PC2"])/(max(datapc[,"PC2"])-min(datapc[,"PC2"]))),
  (max(df[,"PC2"]) - min(df[,"PC1"])/(max(datapc[,"PC1"])-min(datapc[,"PC1"])))
)
datapc <- transform(datapc,
                    v1 = .7 * mult * datapc$PC1,
                    v2 = .7 * mult * datapc$PC2)
# plot
mult <- 3 #multiplier for the arrows and text for envfit
mult.text <- 3
p <- ggplot(scrs) +
  
  # traits loadings
  geom_text(data = datapc, aes(x=v1, y=v2, label = varnames), 
            size=3, vjust=1, color=1) + 
  geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),
               arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color=1) +
  
  # sites
  geom_point(data = scrs, mapping = aes(x = PC1, y = PC2, color = code), alpha = .5) +
  coord_fixed() + ## need aspect ratio of 1!
  
  # envfit
  geom_segment(data = spp.scrs,
               aes(x = 0, xend = mult*PC1, y = 0, yend = mult*PC2),
               arrow = arrow(length = unit(0.2, "cm")), color = 'red') +
  geom_text(data = spp.scrs, 
            aes(x = mult.text*PC1, y = mult.text*PC2, 
                label = Species), 
            size = 4, fontface = "bold", color = "red") +
  xlab("PCoA 1") + ylab("PCoA 2")
p
ggsave(p, filename = "output/stem_pmr_traitord_withCfract.pdf")

```

(c) Only C fractions
```{r}
rhs.Conly
rhs.mat <- pmr.trait.data[,colnames(pmr.trait.data) %in% c(cfract.x.vars)]
rhs.mat.s <- scale(rhs.mat)

# set up all the models
stem.respVars
mod.full.list <- lapply(stem.respVars, function(x) {
  result <- ModelFit_fxn(y = x, y.data = y.mat,
                         rhs = rhs.Conly, rhs.data = rhs.mat.s)
  return(result)
})

#do stepwise model selection
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="backward")
  return(mod.select)
  })

#pull coefs and r2s
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(mod.select.list, unlist(stem.respVars))
fitstat.df<-PullLmFitStats(mod.select.list, unlist(stem.respVars))

fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
r2.df
coefs.df %>%
  left_join(r2.df) -> coefs.df

# which traits were included in the best models?
coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p.stem <- ggplot(tmp, aes(x = respvar, y = term, fill = r.squared)) +
  geom_tile(color = "black")
p.stem

# what direction are the associations?
p.stem <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.stem


p.stem <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.stem
```


(2) by code

Merge trait matrix with the decay parameters estimated by code
```{r}

# decay fits
stemSamples <- load_stemSamples() # this function is in load_microbeData.R
#decayfits <- fit_all_curves(pmr, stemSamples) #this recalculates all the curve fits, uncomment if the data changes; for each species+size
#write_csv(decayfits,"derived_data/decayfits.csv")
decayfits <- read_csv("derived_data/decayfits.csv")
indx <- unique(stemSamples[,c("code","Binomial","size")])
decayfits %>%
  left_join(indx) -> decayfits
code.respVars <- list("alpha","beta","w.t50","w.r2") # weibull params: alpha = shape param, beta = scale param

#merge with trait data, first need to aggregate traits by code (currently at the codeStem level)
trait.data.w.sel %>%
  gather(key = "trait", value = "trait.val", -c(code, size, Stem, codeStem)) %>%
  group_by(code, trait) %>%
  summarize(mean.trait.val = mean(trait.val, na.rm = T)) %>%
  spread(key = "trait", value = "mean.trait.val") -> trait.bycode.w
# now merge
trait.bycode.w %>%
  left_join(decayfits) -> fits.trait.data

dim(fits.trait.data) # only 12 codes/observations


```
Remember there are 12 codes/observations

Set up models
```{r}

orig.x.vars <- c("K","N","P","waterperc")
cfract.x.vars <- c("Arabinose","Galactose","Mannose","perc.Total.lignin","Rhamnose")
code.respVars <- list("alpha","beta","w.t50","w.r2") # weibull params: alpha = shape param, beta = scale param

fits.trait.data %>%
  select(as.character(code.respVars)) -> y.mat
y.mat <- y.mat[,-1]

rhs.orig <- paste(orig.x.vars, collapse = " + ")
rhs.plusC <- paste(c(orig.x.vars, cfract.x.vars), collapse = " + ")
```

(a) Original traits: "K","N","P","waterperc" predict code-level decay parameters
```{r}
# rhs.orig model
rhs.orig <- paste(orig.x.vars, collapse = " + ")
rhs.mat <- fits.trait.data[,colnames(fits.trait.data) %in% orig.x.vars]
rhs.mat.s <- scale(rhs.mat)

# set up all the models
mod.full.list <- lapply(code.respVars, function(x) {
  result <- ModelFit_fxn(y = x, y.data = y.mat,
                         rhs = rhs.orig, rhs.data = rhs.mat.s)
  return(result)
})

#do stepwise model selection
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="backward")
  return(mod.select)
  })

#pull coefs and r2s
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(sum.list, unlist(code.respVars))
fitstat.df<-PullLmFitStats(sum.list, unlist(code.respVars))

fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
coefs.df %>%
  left_join(r2.df) -> coefs.df

# which traits were included in the best models?
coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p.code <- ggplot(tmp, aes(x = respvar, y = term, fill = r.squared)) +
  geom_tile(color = "black")
p.code

# what direction are the associations?
p.code <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.code

```

Visualize code in trait space, add decayfit vectors
```{r}

#create ordination
pca <- prcomp(rhs.mat.s)
summary(pca)

#do envfit
pca.envfit <- envfit(pca, y.mat, na.rm = T)

# make plot
 # collect site scores from mod.obj
scrs <- as.data.frame(scores(pca, display = "sites"))
scrs <- cbind(scrs, code = fits.trait.data$code)
scrs %>%
  left_join(fits.trait.data) -> scrs
  
# collect vector scores from envfit.obj
spp.scrs <- as.data.frame(scores(pca.envfit, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), pval = pca.envfit$vectors$pvals)
  
# collect trait loadings
datapc <- data.frame(varnames=rownames(pca$rotation), pca$rotation[,1:2])
mult <- min(
  (max(df[,"PC1"]) - min(df[,"PC2"])/(max(datapc[,"PC2"])-min(datapc[,"PC2"]))),
  (max(df[,"PC2"]) - min(df[,"PC1"])/(max(datapc[,"PC1"])-min(datapc[,"PC1"])))
)
datapc <- transform(datapc,
                    v1 = .7 * mult * datapc$PC1,
                    v2 = .7 * mult * datapc$PC2)
# plot
mult <- 3 #multiplier for the arrows and text for envfit
mult.text <- 3
p <- ggplot(scrs) +
  
  # traits loadings
  geom_text(data = datapc, aes(x=v1, y=v2, label = varnames), 
            size=3, vjust=1, color=1) + 
  geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),
               arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color=1) +
  
  # sites
  geom_point(data = scrs, mapping = aes(x = PC1, y = PC2, color = code), alpha = .5) +
  coord_fixed() + ## need aspect ratio of 1!
  
  # envfit
  geom_segment(data = spp.scrs,
               aes(x = 0, xend = mult*PC1, y = 0, yend = mult*PC2),
               arrow = arrow(length = unit(0.2, "cm")), color = 'red') +
  geom_text(data = spp.scrs, 
            aes(x = mult.text*PC1, y = mult.text*PC2, 
                label = Species), 
            size = 4, fontface = "bold", color = "red") +
  xlab("PCoA 1") + ylab("PCoA 2")
p
ggsave(p, filename = "output/code_decayfits_traitord.pdf")

```


(b) Original traits plus C fractions...
```{r}
# rhs.orig model
rhs.plusC <- paste(c(orig.x.vars, cfract.x.vars), collapse = " + ")
rhs.mat <- fits.trait.data[,colnames(fits.trait.data) %in% c(orig.x.vars, cfract.x.vars)]
rhs.mat.s <- scale(rhs.mat)

# set up all the models
mod.full.list <- lapply(code.respVars, function(x) {
  result <- ModelFit_fxn(y = x, y.data = y.mat,
                         rhs = rhs.plusC, rhs.data = rhs.mat.s)
  return(result)
})

#do stepwise model selection
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="backward")
  return(mod.select)
  })

#pull coefs and r2s
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(sum.list, unlist(code.respVars))
fitstat.df<-PullLmFitStats(sum.list, unlist(code.respVars))
fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
coefs.df %>%
  left_join(r2.df) -> coefs.df

# which traits were included in the best models?
coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p.code <- ggplot(tmp, aes(x = respvar, y = term, fill = r.squared)) +
  geom_tile(color = "black")
p.code

# what direction are the associations?
p.code <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.code

p.code <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.code

```

Visualize code in trait space, add decayfit vectors
```{r}

#create ordination
pca <- prcomp(rhs.mat.s)
summary(pca)

#do envfit
pca.envfit <- envfit(pca, y.mat, na.rm = T)

# make plot
 # collect site scores from mod.obj
scrs <- as.data.frame(scores(pca, display = "sites"))
scrs <- cbind(scrs, code = fits.trait.data$code)
scrs %>%
  left_join(fits.trait.data) -> scrs
  
# collect vector scores from envfit.obj
spp.scrs <- as.data.frame(scores(pca.envfit, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), pval = pca.envfit$vectors$pvals)
  
# collect trait loadings
datapc <- data.frame(varnames=rownames(pca$rotation), pca$rotation[,1:2])
mult <- min(
  (max(df[,"PC1"]) - min(df[,"PC2"])/(max(datapc[,"PC2"])-min(datapc[,"PC2"]))),
  (max(df[,"PC2"]) - min(df[,"PC1"])/(max(datapc[,"PC1"])-min(datapc[,"PC1"])))
)
datapc <- transform(datapc,
                    v1 = .7 * mult * datapc$PC1,
                    v2 = .7 * mult * datapc$PC2)
# plot
mult <- 3 #multiplier for the arrows and text for envfit
mult.text <- 3
p <- ggplot(scrs) +
  
  # traits loadings
  geom_text(data = datapc, aes(x=v1, y=v2, label = varnames), 
            size=3, vjust=1, color=1) + 
  geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),
               arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color=1) +
  
  # sites
  geom_point(data = scrs, mapping = aes(x = PC1, y = PC2, color = code), alpha = .5) +
  coord_fixed() + ## need aspect ratio of 1!
  
  # envfit
  geom_segment(data = spp.scrs,
               aes(x = 0, xend = mult*PC1, y = 0, yend = mult*PC2),
               arrow = arrow(length = unit(0.2, "cm")), color = 'red') +
  geom_text(data = spp.scrs, 
            aes(x = mult.text*PC1, y = mult.text*PC2, 
                label = Species), 
            size = 4, fontface = "bold", color = "red") +
  xlab("PCoA 1") + ylab("PCoA 2")
p
ggsave(p, filename = "output/code_decayfits_traitord_withCfract.pdf")

```


(a-forward) Original traits but using forward selection instead of backward
```{r}
# rhs.orig model
rhs.orig <- paste(orig.x.vars, collapse = " + ")
rhs.mat <- fits.trait.data[,colnames(fits.trait.data) %in% orig.x.vars]
rhs.mat.s <- scale(rhs.mat)

# set up all the models
mod.full.list <- lapply(code.respVars, function(x) {
  result <- ModelFit_fxn(y = x, y.data = y.mat,
                         rhs = rhs.orig, rhs.data = rhs.mat.s)
  return(result)
})

#do stepwise model selection
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="forward")
  return(mod.select)
  })

#pull coefs and r2s
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(sum.list, unlist(code.respVars))
fitstat.df<-PullLmFitStats(sum.list, unlist(code.respVars))

fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
coefs.df %>%
  left_join(r2.df) -> coefs.df

# which traits were included in the best models?
coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p.code <- ggplot(tmp, aes(x = respvar, y = term, fill = r.squared)) +
  geom_tile(color = "black")
p.code

# what direction are the associations?
p.code <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.code

```

(b-forward) Original traits plus C fractions...
```{r}
# rhs.orig model
rhs.plusC <- paste(c(orig.x.vars, cfract.x.vars), collapse = " + ")
rhs.mat <- fits.trait.data[,colnames(fits.trait.data) %in% c(orig.x.vars, cfract.x.vars)]
rhs.mat.s <- scale(rhs.mat)

# set up all the models
mod.full.list <- lapply(code.respVars, function(x) {
  result <- ModelFit_fxn(y = x, y.data = y.mat,
                         rhs = rhs.plusC, rhs.data = rhs.mat.s)
  return(result)
})

#do stepwise model selection
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="forward")
  return(mod.select)
  })

#pull coefs and r2s
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(sum.list, unlist(code.respVars))
fitstat.df<-PullLmFitStats(sum.list, unlist(code.respVars))
fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
coefs.df %>%
  left_join(r2.df) -> coefs.df

# which traits were included in the best models?
coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p.code <- ggplot(tmp, aes(x = respvar, y = term, fill = r.squared)) +
  geom_tile(color = "black")
p.code

# what direction are the associations?
p.code <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.code

p.code <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.code

```

