---
title: "Does endophyte composition explain decay?"
author: "Marissa Lee"
date: "12/2/2018"
output: github_document
---

```{r, include = F}
#chunk options
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

Load libraries and functions
```{r}

#---------------------------#
#libraries
library(knitr) #r markdown
library(readr) #read/write
library(magrittr) #data formatting
library(tidyr) #data formatting
library(dplyr) #data formatting
library(ggplot2) #plotting
library(grid) #plotting
library(gridExtra)
#devtools::install_github("cornwell-lab-unsw/litterfitter") #fit decay models to timeseries
library(litterfitter) #fit decay models to timeseries
library(lmtest) # for coxtest()
library(vegan) #diversity metrics
library(rioja) #WA-PLS

#---------------------------#
#fxns
source('code/helper_fxns.R')
sourceDir('code')

```

Load decay data, microbial community data, and residuals from "woodTraits_explainDecay.Rmd", and trait data
```{r}
#---------------------------#
#data

# stem mass
initial_mass <- read_in_initial_mass()
harvest_mass <- LoadHarvestFiles()
pmr <- Calc_massRemaining(initial_mass, harvest_mass) # calculate percent mass remaining (pmr)
pmr_byStem <- AvgPMR_byStem(pmr) # average pmr for each stem and timepoint
stem.respVars <- list("time7", "time13", "time25", "time37","time59")

# decay fits
stemSamples <- load_stemSamples() # this function is in load_microbeData.R
#decayfits <- fit_all_curves(pmr, stemSamples) #this recalculates all the curve fits, uncomment if the data change; for each species+size
#write_csv(decayfits,"derived_data/decayfits.csv")
decayfits <- read_csv("derived_data/decayfits.csv")
indx <- unique(stemSamples[,c("code","Binomial","size")])
decayfits %>%
  left_join(indx) -> decayfits
code.respVars <- list("w.t50","alpha","beta","w.r2") # weibull params: alpha = shape param, beta = scale param

# microbes
microb.data <- load_MicrobeCollection(stemSamples)
taxAndFunguild <- microb.data$taxAndFunguild
comm.otu <- microb.data$comm.otu
seqSamples <- microb.data$seqSamples

# residuals from 'woodTraits_explainDecay.Rmd'
resids_explainDecay.list <- readRDS(file = "derived_data/resids_explainDecay_list.RData")

# wood traits
traits.code <- trait.means_byCode(stemSamples, fill.densitybark = TRUE) # averaged by code; if fill.densitybark == TRUE, then use small stem estimates to approximate large stem density and bark thickness
traits.stem <- trait.means_byStem(stemSamples) # averaged by codeStem
trait.order <- traitcol.order()
traitVars <- c(trait.order$phys.cols, trait.order$nutr.cols)
traitVars.stem <- trait.order$stem.trait.cols
traitVars.cfract <- trait.order$cfract.cols

```

Figure S1. Sample-effort curves
```{r}
#this takes a while... uncomment if the data changes
#plot_sampleEffortCurves(comm.otu) # output/figures/supplementary/sampleEffortCurve.pdf

```

Filter community matrix to include only taxa that are present in a least 20% of all the samples. This step removes taxa that may not contribute much to our understanding of the relationship between species’ multivariate abundance and environment.
```{r}
comm.otu.trimmed <- removeRareOTUs(comm.otu)
```

Initalize lists to store results
```{r}
# save the objects that come out of makefig__...
models_explainDecay.list <- list() 

# save the plots that dig into models summary results
plots_explainDecay.list <- list()

```


##########################################
# Community predicts decay

## *Hyp (species+size-level)* Species+size-level (average) initial microbial community composition will predict variation in decay model fit (r2), rate (t70, k), and lagginess (alpha).
```{r, message = F, results = F}
cvfit.results.code <- doAnalysis_endoComp_explainDecay(comm.otu, seqSamples, decayfits, code.respVars,
                                                       use.cache = T, save.cache = T)
cvfit.results.code$prettyTabs # will merge this with the other summary table to export as a csv in the next chunk

models_explainDecay.list[['comm-decay.code']] <- cvfit.results.code$prettyTabs
```

None of the components are significant.


## *Hyp (stem-level)* Stem-level initial microbial communitiy compositions will predict variation in percent mass loss, particularly in the early stages of decay.
```{r, message = F, results = F}

cvfit.results.stem <- doAnalysis_endoComp_explainPMR(comm.otu, pmr_byStem, stem.respVars)
cvfit.results.stem$prettyTabs

models_explainDecay.list[['comm-decay.stem']] <- cvfit.results.stem$prettyTabs
```

Comp01 (of the non-trimmed community) is a significant predictor of percent mass remaining at 37 months. 


Plot the distribution of WA-PLS scores. Just out of curiousity, I'd like to pull in OTU "niche" info from the boral analysis to see if there's a relationship between OTU WA-PLS scores and wood trait coeffient estimates. There's a weak negative relationship between an OTU's WA-PLS score and waterperc coefficient (slope=-2.3, p=.03), suggesting that OTUs that "prefer" high-water niche space are associated with less mass remaining at time37.
```{r}
plots_explainDecay.list[['comm-decay.stem']] <- makefig__wapls_score_time37(cvfit.results.stem,
                                                                            taxAndFunguild)
grid.arrange(plots_explainDecay.list[['comm-decay.stem']]$p1, 
             plots_explainDecay.list[['comm-decay.stem']]$p2, ncol = 2)

```

Is this because there is an underlying signature of wood traits on the initial microbial community that is driving the relationship between the community and the mass remaining after 37 months?  The next analysis (Community predicts decay residuals) will test this formally.



##########################################
# Community predicts decay residuals (already explained by trait/Cfraction)

# Slice 1: Exclude C fraction data

## *Hyp (species+size-level)* After accounting for variation in decay due to wood traits, average initial microbial communitiy compositions will predict variation in decay model fit (r2), rate (t70, k), and lagginess (alpha).
```{r, message=FALSE, results = F}
resids_explainDecay.list$traits.code


cvfit.results.code <- doAnalysis_endoComp_explainDecayResids(comm.otu, seqSamples, decayfits, 
                                                             code.respVars, 
                                                             traitResiduals.code = resids_explainDecay.list$traits.code,
                                                             use.cache = T, save.cache = T)
cvfit.results.code$prettyTabs
cvfit.results.code

models_explainDecay.list[['comm-decayResidtraits.code']] <- cvfit.results.code$prettyTabs

```

## *Hyp (species+size-level)* After accounting for variation in decay due to C fractions, average initial microbial communitiy compositions will predict variation in decay model fit (r2), rate (t70, k), and lagginess (alpha).
```{r, message=FALSE, results = F}
resids_explainDecay.list$cfract.code

cvfit.results.code <- doAnalysis_endoComp_explainDecayResids(comm.otu, seqSamples, decayfits, 
                                                             code.respVars, 
                                                             traitResiduals.code = resids_explainDecay.list$cfract.code,
                                                             use.cache = T, save.cache = T)
cvfit.results.code$prettyTabs
cvfit.results.code

models_explainDecay.list[['comm-decayResidtraits.code']] <- cvfit.results.code$prettyTabs

```


Community data doesn't improve our understanding of decay rates (k, t70) or variation in decay rate (ne.r2) beyond what is known from the trait data.


## *Hyp (stem-level)* After accounting for variation in decay due to wood traits (no models with density, includes small-species level bark thickness), stem-specific initial microbial communitiy compositions will predict variation in percent mass loss, particularly in the early stages of decay.
```{r, message=FALSE, results = F}
cvfit.results.stem <- doAnalysis_endoComp_explainPMRResids(comm.otu, 
                                                           pmr_byStem, 
                                                           stem.respVars, 
                                                           traitResiduals.stem =
                                                             resids_explainDecay.list$traits.stem)
cvfit.results.stem$prettyTabs

models_explainDecay.list[['comm-decayResidtraits.stem']] <- cvfit.results.stem$prettyTabs

#Notes about how to interpret WA-PLS...
#rand.t.test(fit.tr.t13.cv.nt) # comp2 has a signif pval, but...
#screeplot(fit.tr.t13.cv.nt)
#There are two cases where there's a large % decrease in model RMSE from Component 1 to Component 2. This happens when using the whole community dataset to predict trait residuals at time13 and time37. In both cases the cross-validated RMSE is way higher than the model RMSE for all the components, suggesting that even the first community component doesn't perform well on the leave-one-out validation dataset.  Also, the cross-validated R-squared values (correlation between the observed and predicted values from the "loo" validation dataset) show that the model fit decreases after Component 1.  If there were a global maximum such that we saw an increase in R2 after adding more Components then maybe we could interpret Component 2, but there is no evidence of a better fitting model with more components based on the cross-validation results.



```

Community data doesn't improve our understanding of mass loss (pmr after 7, 13, 25, 37, and 59 months) beyond what is known from the trait data.



# Slice 2: Include C fraction data (and exclude small stems)

## *Hyp (species+size-level)* After accounting for variation in decay due to wood C fractions, average initial microbial communitiy compositions will predict variation in decay model fit (r2), rate (t70, k), and lagginess (alpha).
```{r, message=FALSE, results = F}
cvfit.results.code <- doAnalysis_endoComp_explainDecayResids(comm.otu, seqSamples, decayfits, 
                                                             code.respVars, 
                                                             traitResiduals.code = resids_explainDecay.list$cfract.code,
                                                             use.cache = T, save.cache = T)
cvfit.results.code$prettyTabs

models_explainDecay.list[['comm-decayResidcfract.code']] <- cvfit.results.code$prettyTabs

```

Community data doesn't improve our understanding of decay rates (k, t70) or variation in decay rate (ne.r2) beyond what is known from the C fraction data.


## *Hyp (stem-level)* After accounting for variation in decay due to wood C fractions (no models with density, includes small-species level bark thickness), stem-specific initial microbial communitiy compositions will predict variation in percent mass loss, particularly in the early stages of decay.
```{r, message=FALSE, results = F}
cvfit.results.stem <- doAnalysis_endoComp_explainPMRResids(comm.otu, pmr_byStem, 
                                                           stem.respVars, 
                                                           traitResiduals.stem =
                                                             resids_explainDecay.list$cfract.stem)
cvfit.results.stem$prettyTabs

models_explainDecay.list[['comm-decayResidcfract.stem']] <- cvfit.results.stem$prettyTabs

```

Community data doesn't improve our understanding of mass loss (pmr after 7, 13, 25, 37, and 59 months) beyond what is known from the C fraction data.


##########################################
# Relationship between wood traits and community --- GET RID OF THIS SECTION? Only look for signiture of traits in community comp if there are relationships between comp and decay.. there were none

# Slice 1: Exclude C fraction data

## *Hyp (species+size-level)* Initial microbial communitiy compositions will covary with initial wood traits

Use ordistep to find the best combination of traits to constrain variation in the endophyte community
```{r}
# mods <- doAnalysis_endoComp_woodTraits(comm.otu, seqSamples, traits.code, traitVars, 
#                                        use.cache = T, save.cache = F,
#                                        use.cache.ordi = T, save.cache.ordi = F,
#                                        cfract = F)
# 
# # trimmed communities only 
# curr.mod <- mods$mod.t.code
# curr.data <- mods$datasets.t.code
# 
# # identify the predictor variables
# tmp <- rownames(anova(curr.mod, by = "margin"))
# xvars <- tmp[tmp != "Residual"]
# xvars
# curr.data$traits %>%
#   select(xvars) -> xvars.mat
# 
# # total proportion constrained
# extract_constrainedInertia_proport(curr.mod)
# 
# # do the varpart
# part <- varpart(curr.data$comm, ~size, ~barkthick, ~ waterperc, data = xvars.mat)
# part.df <- data.frame(part$part$fract[1:3,], xvars)
# part.df

# makefigs__endoComp_woodTraits(comm.otu, seqSamples, traits.code) # output/figures/supplementary/dbRDA_code.pdf
# tab.list <- maketabs__endoComp_woodTraits(comm.otu, seqSamples, traits.code, use.cache = T)
# tab.list
# write.csv(tab.list$an.nt.code, file="output/tables/supplementary/dbRDAanova_nt_code.csv")
# write.csv(tab.list$an.t.code, file="output/tables/supplementary/dbRDAanova_t_code.csv")


```

## *Hyp (stem-level)* Average initial microbial communitiy compositions will covary with initial wood traits
```{r}

# mods <- doAnalysis_endoComp_woodTraits.stem(comm.otu, traits.code, traits.stem)
# 
# # trimmed communities only 
# curr.mod <- mods$mod.t.stem
# curr.mod

```


# Slice 2: Include C fraction data (and exclude small stems)

## *Hyp (species+size-level)* Initial microbial communitiy compositions will covary with initial wood traits

Use ordistep to find the best combination of traits to constrain variation in the endophyte community
```{r}
# traitVars.cfract
# mods <- doAnalysis_endoComp_woodTraits(comm.otu, seqSamples, traits.code, traitVars = traitVars.cfract, 
#                                        use.cache = T, save.cache = F,
#                                        use.cache.ordi = F, save.cache.ordi = F,
#                                        cfract = T) # don't cache it will write over the one for traits
# 
# # trimmed communities only
# curr.mod <- mods$mod.t.code
# curr.data <- mods$datasets.t.code
# 
# # identify the predictor variables
# curr.mod # none

```


## *Hyp (stem-level)* Average initial microbial communitiy compositions will covary with initial wood traits
```{r}
#makefigs__endoComp_woodTraits.stem(comm.otu, traits.code, traits.stem) # output/figures/supplementary/dbRDA_stem.pdf

```


##########################################
# Print results

```{r}

# save the objects that come out of makefig__...
names(models_explainDecay.list)
filenames<- paste0("output/tables/supplementary/wapls_", names(models_explainDecay.list),".csv")
for(i in 1:length(models_explainDecay.list)){
  write.csv(models_explainDecay.list[[i]], file = filenames[[i]])
}

# save the plots that dig into models summary results
pdf(file = "output/figures/supplementary/wapls_pmr37.pdf")
grid.arrange(plots_explainDecay.list$`comm-decay.stem`$p1,
             plots_explainDecay.list$`comm-decay.stem`$p2, ncol = 2)
dev.off()

```

Show the relationship between pmr time37, sample water, and wapls
```{r}
samp.score <- cvfit.results.stem$cvfit.notrim$time37$T[,'Comp01']
wapls.df <- data.frame(codeStem = names(samp.score), wapls= samp.score)
wapls.df %>%
  left_join(traits.stem) %>%
  select(codeStem, wapls, waterperc) %>%
  left_join(pmr_byStem) %>%
  select(codeStem, size, wapls, waterperc, time37) -> df
  
p1 <- ggplot(df, aes(x = wapls, y = time37, color = size)) +
  geom_point() +
  ylab("Percent mass remaining after 3 yrs") + 
  xlab("Endophyte Component 1") +
  theme_classic() +
  scale_color_manual(values = c("darkgray","black")) +
  guides(color = F)
p1
p2 <- ggplot(df, aes(x = wapls, y = waterperc, color = size)) +
  geom_point() +
  ylab("Wood water (%)") + 
  xlab("Endophyte Component 1") +
  theme_classic() +
  scale_color_manual(values = c("darkgray","black")) +
  guides(color = F)
p2
p3 <- ggplot(df, aes(x = waterperc, y = time37, color = size)) +
  geom_point() +
  ylab("Percent mass remaining after 3 yrs") + 
  xlab("Wood water (%)") +
  theme_classic()+
  scale_color_manual(values = c("darkgray","black")) +
  guides(color = F)
p3

pdf(file = "output/figures/maintext/water-comm-pmr37.pdf")
grid.arrange(p1, p3, p2, ncol = 2)
dev.off()


```

