---
title: "Does chemistry or community better predict mass loss?"
author: "Marissa Lee"
date: "10/23/2017"
output: github_document
---

```{r,message=FALSE}
#chunk options
knitr::opts_chunk$set(echo = TRUE)

#libraries
devtools::install_github("cornwell-lab-unsw/litterfitter")
library(dplyr)
library(ggplot2)
library(readr)
library(vegan)
library(litterfitter)
library(magrittr)
library(tidyr)
source("code/load_fxns.R")
source("code/curve_fitting_fxns.R")
source("code/distance_fxns.R")
library(gridExtra)
```

### Load microbial community data
```{r }
fung.otu<-load_matotu()
comm.otu<-add_oomycetes(fung.otu)
#plot_sampleEffortCurves(comm.otu)
```

### Load wood trait data
```{r}
trait.means<-mergeTraitData()
traits.long<-as.data.frame(gather(trait.means, key=trait, value=value, -(1:3)))

ggplot(traits.long, aes(x=species_lower, y=value, color=size)) + 
  geom_point() + 
  facet_wrap(~trait, scales="free") +
  mytheme + theme(axis.text.x = element_text(angle = 90, hjust = 1))

# #the Fe values for large stems are not 0, they are just small values
# subset(traits.long, trait=="Fe") %>%
#   ggplot(aes(x=species_lower, y=value, color=size)) + 
#   geom_point() +
#   facet_wrap(~size, scales="free")
```

### Load mass loss data
```{r}

#initial mass
initial_mass <- read_in_initial_mass()
# initial_mass %>%
#   mutate(SpeciesCode=tolower(Species))%>%
#   ggplot(aes(y=totalSampleDryMass,x=SpeciesCode,fill=size))+
#   geom_violin()+ 
#   theme(axis.text.x=element_text(angle=90,hjust=1)) + 
#   scale_y_log10()

#mass at harvest
harvest_mass<-LoadHarvestFiles()

#create a complete sample mass df for all time points
mass.data<-rbind(initial_mass, harvest_mass)

```
Identify outliers in mass loss data
```{r}
mass.data %>% 
  ggplot(aes(x=time, y=totalSampleDryMass)) + geom_point(alpha=0.6)+theme_bw() 
mass.data[which.max(mass.data$totalSampleDryMass),]
outlier.uniques<-as.character(mass.data[which.max(mass.data$totalSampleDryMass),"unique"])
outlier.uniques
```
...another view...might want to check out those two high mass value outliers from harvest 3
```{r}
mass.data %>%
  ggplot(aes(x=time, y=totalSampleDryMass,col=size)) +
  geom_point(position="jitter",alpha=0.6)+theme_bw()+scale_y_log10()

#Might want to check out those two high mass value outliers from harvest 3
filter(mass.data, size=="small", time==25) %>%
  arrange(-totalSampleDryMass)->tmp
tmp[1:2,]
outlier.uniques<-c(outlier.uniques, tmp$unique[1:2])
```
Are these real 0's?
```{r}
mass.data[which(mass.data$totalSampleDryMass==0),]
```
Remove outliers from mass.data and merge time zero with the other harvests to calculate proportion mass remaining at each time point
```{r}
mass.data<-mass.data[!mass.data$unique %in% outlier.uniques,]

#Merge time zero with the other harvests to calculate proportion mass remaining at each time point
mass.data %>%
  filter(time==0) %>%
  rename(timeZeroDensity=density) %>%
  rename(timeZeroMass=totalSampleDryMass) %>%
  select(unique,timeZeroMass,timeZeroDensity)->time_zero

mass.data %>%
  left_join(time_zero,by="unique") %>%
  mutate(pmr=totalSampleDryMass/timeZeroMass) %>%
  mutate(SpeciesCode=tolower(Species)) -> plotting_df

```


### Non-linear curve fits of decay trajectories
Using `litterfitter` to apply both negative exponenial and weibull to all species/size classes
```{r,warning=FALSE,message=FALSE,results=FALSE}

#spdf <- fit_all_curves(plotting_df) #this recalculates all the curve fits, uncomment if the data changes
spdf <- read_csv("derived_data/mass_loss_parameters.csv")
#write_csv(spdf,"derived_data/mass_loss_parameters.csv")


ggplot(spdf,aes(x=t70,y=w.t70,col=size))+geom_point()+labs(x="Time to 30% mass loss (negative exponential)", y="weibull time to 30% mass loss")+geom_abline(slope=1,intercept=0,linetype="dashed")+theme_bw()

# #Here is an example where the neg.exp and the weibull curves are almost identical
# plotting_df %>%
#   filter(SpeciesCode=="eute",size=="small")  ->one_example
# plot_multiple_fits(time = one_example$time/12,
#                    mass.remaining = one_example$pmr,
#                    bty = 'n', model = c('neg.exp', 'weibull'),
#                    xlab = 'Time', ylab = 'Proportion mass remaining',iters=1000)
# #and one where they are pretty different:
# plotting_df %>%
#   filter(SpeciesCode=="ripi",size=="small") -> another_example
# plot_multiple_fits(time = another_example$time/12,
#                    mass.remaining = another_example$pmr,
#                    bty = 'n', model = c('neg.exp', 'weibull'),
#                    xlab = 'Time', ylab = 'Proportion mass remaining',iters=1000)
```


### Testing effects of t=0 points
replotting the last example without the t=0 points
```{r}
#why filtering for ripi small?

plotting_df %>%
  filter(SpeciesCode=="ripi",size=="small",time>0) ->out

plot_multiple_fits(time = out$time/12,
                   mass.remaining = out$pmr,
                   bty = 'n', model = c('neg.exp', 'weibull'),
                   xlab = 'Time', ylab = 'Proportion mass remaining',iters=1000)
```
Checking that the fits are the same for weibull which they are 
```{r}

out%$%
fit_litter(time = time/12, 
        mass.remaining = pmr, model = c("weibull"), iters = 1000) ->plot_1
print(plot_1)
plot(plot_1)

#why filtering for ripi small?

plotting_df %>%
  filter(SpeciesCode=="ripi",size=="small") ->out
out%$%
fit_litter(time = time/12, 
        mass.remaining = pmr, model = c("weibull"), iters = 1000) ->plot_2
print(plot_2)
plot(plot_2)
```
Conclusion: including t=0 points affects the liklihood and the model selection criteria, but the curve fits are identical with this formulation.  Excluding the t=0 fits has an effect of prefering simpler models, which is the same effect as increasing the penalty for model complexity.  

### Plot beta diversity of microbial community vs distance in decay params
```{r}

#calculate pairwise community distances
sampTab<-CreateSeqSampTab(mass.data) #1. creat sample tab for community sequences samples
comm.dist<-Calc_commDists(sampTab, comm.otu) #2. calc the distances
summ.comm_dist<-SummarizeCommDist_byCodePair(comm.dist) #3. summarize the distances by codePair

#calculate pairwise decay parameter distances
spdf<-mutate(spdf, minAIC = pmin(neg.exp.aic, w.aic) ) #1. find the min AIC for each species+size
spdf<-AddCodeID(sampTab) #2. add code to spdf using sampTab
aic.dist<-Calc_decayParamDiffs(valueCol="minAIC", spdf, sampTab) #3. calc decay param diffs, e.g. minAIC
k.dist<-Calc_decayParamDiffs(valueCol="k", spdf, sampTab)
alpha.dist<-Calc_decayParamDiffs(valueCol="alpha", spdf, sampTab)

#merge community and decay param distances into the same dataframe
join.dist.aic<-MergeCommNDecaydists_byCodePair(decayparam.dist=aic.dist, summ.comm_dist)
join.dist.k<-MergeCommNDecaydists_byCodePair(decayparam.dist=k.dist, summ.comm_dist)
join.dist.alpha<-MergeCommNDecaydists_byCodePair(decayparam.dist=alpha.dist, summ.comm_dist)

# delta minAIC
p.aic<-ggplot(join.dist.aic, aes(x=mean_comm_dist, y=abs(decayparam_dist), color=size)) + 
  geom_point() + 
  geom_errorbarh(aes(xmin=lower_comm_dist, xmax=upper_comm_dist)) +
  xlab('Mean microbial community distance') + ylab("Delta decay model AIC")

# delta k
p.k<-ggplot(join.dist.k, aes(x=mean_comm_dist, y=abs(decayparam_dist), color=size)) + 
  geom_point() + 
  geom_errorbarh(aes(xmin=lower_comm_dist, xmax=upper_comm_dist)) +
  xlab('Mean microbial community distance') + ylab("Delta decay model k")

# delta alpha
p.alpha<-ggplot(join.dist.alpha, aes(x=mean_comm_dist, y=abs(decayparam_dist), color=size)) + 
  geom_point() + 
  geom_errorbarh(aes(xmin=lower_comm_dist, xmax=upper_comm_dist)) +
  xlab('Mean microbial community distance') + ylab("Delta decay model alpha")

grid.arrange(p.aic, p.k, p.alpha)


#combine delta decay param columns
j1<-rename(join.dist.aic, "aic_dist"="decayparam_dist") %>%
  select(codePair, size, mean_comm_dist, aic_dist)
j2<-rename(join.dist.k, "k_dist"="decayparam_dist") %>%
  select(codePair, size, mean_comm_dist, k_dist)
j3<-rename(join.dist.alpha, "alpha_dist"="decayparam_dist") %>%
  select(codePair, size, mean_comm_dist, alpha_dist)

left_join(j1, j2) %>%
  left_join(j3) -> join.dist

join.dist<-mutate(join.dist, aic_dist=abs(aic_dist)) %>%
  mutate(k_dist=abs(k_dist)) %>%
  mutate(alpha_dist=abs(alpha_dist))


```

### Plot beta diversity of wood traits vs distance in decay params
```{r}

#calculate pairwise wood trait distances




```



