---
title: "Does chemistry or community better predict mass loss?"
author: "Marissa Lee"
date: "10/23/2017"
output: github_document
---

```{r,message=FALSE}
#chunk options
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

#libraries
devtools::install_github("cornwell-lab-unsw/litterfitter")
library(dplyr)
library(ggplot2)
library(readr)
library(vegan)
library(knitr)
library(litterfitter)
library(magrittr)
library(tidyr)
library(gridExtra)
library(rioja)

#fxns
source("code/load_fxns.R")
source("code/curve_fitting_fxns.R")
source("code/distance_fxns.R")
source("code/otuIDs_fxns.R")


### LOAD MICROBIAL COMMUNITY DATA

#stem sample meta data
#stemSamples<-load_stemSamples() #uncomment if the data changes
#write_csv(stemSamples, "derived_data/stemSamples.csv")
stemSamples<-read_csv("derived_data/stemSamples.csv")

#OTU table
#fung.otu<-load_matotu() #uncomment if the data changes
#comm.otu<-add_oomycetes(fung.otu) #add the oomycetes #uncomment if the data changes
#write.csv(comm.otu, "derived_data/comm_otu.csv")
comm.otu<-read.csv("derived_data/comm_otu.csv", row.names=1)

#create sequence sample meta data table
#seqSamples<-load_seqSamples(comm.otu, stemSamples) #uncomment if the data changes
#write_csv(seqSamples, "derived_data/seqSamples.csv")
seqSamples<-read_csv("derived_data/seqSamples.csv")

#taxon lookup info
#taxAndFunguild<-load_TaxAndFunguild(comm.otu) #uncomment if the data changes
#write_csv(taxAndFunguild, "derived_data/taxaAndFunguild.csv")
taxAndFunguild<-read_csv("derived_data/taxaAndFunguild.csv")

#plot_sampleEffortCurves(comm.otu)


### LOAD WOOD TRAIT DATA
#traits.mean<-mergeTraitData() #uncomment if the data changes
#write_csv(traits.mean, "derived_data/traits_mean.csv") 
traits.mean<-read_csv("derived_data/traits_mean.csv")
#missing data
#traits.long<-as.data.frame(gather(traits.mean, key=trait, value=value, -(1:3)))
#filter(traits.long, is.na(value))


### LOAD MASS LOSS DATA and CALCULATE % MASS REMAINING AT EACH TIMEPOINT
#initial_mass <- read_in_initial_mass() #uncomment if the data changes
#harvest_mass<-LoadHarvestFiles()
#mass.data<-bind_rows(initial_mass, harvest_mass)
#missing data
#mass.data %>% filter(is.na(totalSampleDryMass))
#plotting_df<-Calc_massRemaining(mass.data)
#matching failures
# plotting_df %>%
#   filter(is.na(pmr)) %>%
#   select(unique, species, size, time, totalSampleDryMass, notes) %>%
#   spread(key=time, value=totalSampleDryMass)
#remove NAs
#plotting_df %>% filter(!is.na(pmr)) -> plotting_df
#write_csv(plotting_df,"derived_data/plotting_df.csv")
plotting_df<-read_csv("derived_data/plotting_df.csv")

### CALCULATE DECAY TRAJECTORY FITS
#spdf <- fit_all_curves(plotting_df) #this recalculates all the curve fits, uncomment if the data changes
#indx<-select(stemSamples, code, species, size)
#spdf<-left_join(spdf, indx) #add code
#write_csv(spdf,"derived_data/mass_loss_parameters.csv")
spdf <- read_csv("derived_data/mass_loss_parameters.csv")
# ggplot(spdf,aes(x=t70,y=w.t70,col=size))+
#   geom_point()+
#   labs(x="Time to 30% mass loss (negative exponential)", 
#        y="Time to 30% mass loss (Weibull)")+
#   geom_abline(slope=1,intercept=0,linetype="dashed")+theme_bw()

```



1. Wood traits as a preditor
*Hyp:* Variation in wood traits will lead to differences in decay model fit (r2), rate (k), and lagginess (alpha).  Specifically, we expect samples with (a) high waterperc, (b) low density and C, (c) high P, K, Ca, Mn, Fe, Zn, and N, and (d) thicker bark (potential mech: limiting microbial colonization) to have better-fiting decay models (r2), faster decay rates (k), and less lagginess (alpha). 
```{r}

#merge into 1 df
spdf %>%
  select(code, species, size, ne.r2, k, t70, alpha) %>%
  left_join(traits.mean) %>% 
  filter(!is.na(P)) %>%
  filter(!is.na(waterperc)) -> spdf.traits

#fit full models
mod.full.r<-lm(ne.r2 ~ size + waterperc + density + barkthick + P + K + Ca + Mn + Fe + Zn + N + C, data=spdf.traits)
mod.full.k<-lm(k ~ size + waterperc + density + barkthick + P + K + Ca + Mn + Fe + Zn + N + C, data=spdf.traits)
mod.full.t70<-lm(t70 ~ size + waterperc + density + barkthick + P + K + Ca + Mn + Fe + Zn + N + C, data=spdf.traits)
mod.full.alpha<-lm(alpha ~ size + waterperc + density + barkthick + P + K + Ca + Mn + Fe + Zn + N + C, data=spdf.traits)

#do stepwise model selection
mod.select.r<-step(mod.full.r, direction="backward")
mod.select.k<-step(mod.full.k, direction="backward")
mod.select.t70<-step(mod.full.t70, direction="backward")
mod.select.alpha<-step(mod.full.alpha, direction="backward")

#save the residuals
trait.residuals<-data.frame(code=spdf.traits$code,
  r2.resid=mod.select.r$residuals,
  k.resid=mod.select.k$residuals,
  t70.resid=mod.select.t70$residuals,
  alpha.resid=mod.select.alpha$residuals)

```

*Result:*
- r2
```{r}
summary(mod.select.r) # waterperc, C
#ggplot(spdf.traits, aes(x=waterperc, y=ne.r2, color=species, shape=size)) + geom_point()
```
- k
```{r}
summary(mod.select.k) # size, waterperc, barkthick, Ca, Zn, N
ggplot(spdf.traits, aes(x=waterperc, y=k, color=species, size=density)) + geom_point() + facet_grid(~size)
#ggplot(spdf.traits, aes(x=waterperc, y=density, color=species, size=k)) + geom_point() + facet_grid(~size)
```
- t70
```{r}
summary(mod.select.t70) # size, waterperc, barkthick, Ca, Zn
ggplot(spdf.traits, aes(x=waterperc, y=t70, color=species, size=density)) + geom_point() + facet_grid(~size)
```
- alpha
```{r}
summary(mod.select.alpha) # density, Zn, C
```




2. Community as a predictor
*Hyp:* Average initial microbial communitiy compositions will predict variation in decay model fit (r2), rate (k), and lagginess (alpha).  
```{r, include=FALSE}

#average OTU abundances by code
CODE<-unique(seqSamples$code)
meanOTUabund.list<-list()
sdOTUabund.list<-list()
for(i in 1:length(CODE)){
  #identify the seq_sampNames to aggregate
  curr.seqSamps<-seqSamples[seqSamples$code==CODE[i],]$seq_sampName 
  #filter 
  curr.comm.otu<-comm.otu[row.names(comm.otu) %in% curr.seqSamps,]
  #summarize
  meanOTUabund.list[[i]]<-apply(curr.comm.otu, 2, mean, na.rm=TRUE)
  sdOTUabund.list[[i]]<-apply(curr.comm.otu, 2, sd, na.rm=TRUE)
}
names(meanOTUabund.list)<-CODE
names(sdOTUabund.list)<-CODE
meanOTUabund.list %>% combine() %>% as.data.frame %>% t() -> 
  meanOTUabund
sdOTUabund.list %>% combine() %>% as.data.frame %>% t() -> 
  sdOTUabund

#get rid of empty OTU cols
#sum(colSums(meanOTUabund)!=0)
meanOTUabund.trim<-meanOTUabund[,colSums(meanOTUabund)!=0]

#get rid of spdf rows for which there is missing community data... this is eusc
#spdf[!spdf$code %in% row.names(meanOTUabund.trim),]$code
spdf.trim<-spdf[spdf$code %in% row.names(meanOTUabund.trim),]

#fit models
fit.r2 <- WAPLS(meanOTUabund.trim, spdf.trim$ne.r2)
fit.k <- WAPLS(meanOTUabund.trim, spdf.trim$k)
fit.t70 <- WAPLS(meanOTUabund.trim, spdf.trim$t70)
fit.alpha <- WAPLS(meanOTUabund.trim, spdf.trim$alpha)

#cross-validate models using the leave-one-out method
fit.r2.cv <- crossval(fit.r2, cv.method="loo")
fit.k.cv <- crossval(fit.k, cv.method="loo")
fit.t70.cv <- crossval(fit.t70, cv.method="loo")
fit.alpha.cv <- crossval(fit.alpha, cv.method="loo")

#rand.t.test(fit.alpha.cv) #perform randomization t-test to test the significance of a cross-validated model
#screeplot(fit.r2.cv)

```
*Result:*
-r2
```{r}
rand.t.test(fit.r2.cv)
```
-k
```{r}
rand.t.test(fit.k.cv)
```
-t70
```{r}
rand.t.test(fit.t70.cv)
```
-alpha
```{r}
rand.t.test(fit.alpha.cv)
```


3. Community+traits as a predictor
```{r}

#run rioja::WAPLS on wood trait residuals

#make sure that the residuals and community dataframes are aligned
trait.residuals$code<-as.character(trait.residuals$code)
trait.residuals[!trait.residuals$code %in% row.names(meanOTUabund.trim),"code"] #missing eusc from meanOTUabund.trim
row.names(meanOTUabund.trim)[!row.names(meanOTUabund.trim) %in% trait.residuals$code] #missing EUSC and olst from trait.residuals

trait.residuals.trim<-trait.residuals[trait.residuals$code %in% row.names(meanOTUabund.trim),] #trim
meanOTUabund.trim2<-meanOTUabund.trim[row.names(meanOTUabund.trim) %in% trait.residuals.trim$code,] #trim

ord<-match(row.names(meanOTUabund.trim2), trait.residuals.trim$code)
trait.residuals.trim.o<-trait.residuals.trim[ord,]

sum(trait.residuals.trim.o$code != row.names(meanOTUabund.trim2)) #this needs to by 0

#get rid of empty OTU cols
sum(colSums(meanOTUabund.trim2)!=0)
meanOTUabund.trim2<-meanOTUabund.trim2[,colSums(meanOTUabund.trim2)!=0]


#fit models
fit.tr.r2 <- WAPLS(meanOTUabund.trim2, trait.residuals.trim.o$r2.resid)
fit.tr.k <- WAPLS(meanOTUabund.trim2, trait.residuals.trim.o$k.resid)
fit.tr.t70 <- WAPLS(meanOTUabund.trim2, trait.residuals.trim.o$t70.resid)
fit.tr.alpha <- WAPLS(meanOTUabund.trim2, trait.residuals.trim.o$alpha.resid)

#cross-validate models using the leave-one-out method
fit.tr.r2.cv <- crossval(fit.tr.r2, cv.method="loo")
fit.tr.k.cv <- crossval(fit.tr.k, cv.method="loo")
fit.tr.t70.cv <- crossval(fit.tr.t70, cv.method="loo")
fit.tr.alpha.cv <- crossval(fit.tr.alpha, cv.method="loo")

#rand.t.test(fit.alpha.cv) #perform randomization t-test to test the significance of a cross-validated model
#screeplot(fit.r2.cv)

rand.t.test(fit.tr.r2.cv)
rand.t.test(fit.tr.k.cv)
rand.t.test(fit.tr.t70.cv) 
rand.t.test(fit.tr.alpha.cv)


```

4. Diversity (and diversity of specific clades) as a predictor

Plot richness of key players vs decay param distances BETWEEN species+size
```{r}

# summarize the presence of ... in each sample
sapro.df<-Calc_richOTUtype(colNam="Trophic.Mode", grepTerm="Sapro", taxAndFunguild, comm.otu)  
sapList<-Plot_richOTUtype(otutype.df=sapro.df, 
                        valueCol_vec=c("ne.r2", "k","alpha"), 
                        otutypeNam="Saprotroph", spdf)

basidio.df<-Calc_richOTUtype(colNam="phylum", grepTerm="Basid", taxAndFunguild, comm.otu)
basidList<-Plot_richOTUtype(otutype.df=basidio.df, 
                        valueCol_vec=c("ne.r2", "k","alpha"), 
                        otutypeNam="Basidio", spdf)

path.df<-Calc_richOTUtype(colNam="Trophic.Mode", grepTerm="Patho", taxAndFunguild, comm.otu)
pathList<-Plot_richOTUtype(otutype.df=path.df, 
                        valueCol_vec=c("ne.r2", "k","alpha"), 
                        otutypeNam="Pathogen", spdf)

oomy.df<-Calc_richOTUtype(colNam="kingdom", grepTerm="Protist", taxAndFunguild, comm.otu)
ooList<-Plot_richOTUtype(otutype.df=oomy.df, 
                        valueCol_vec=c("ne.r2", "k","alpha"), 
                        otutypeNam="Oomycete", spdf)

#plot
grid.arrange(sapList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             sapList[['k']] + guides(color=FALSE, shape=FALSE), 
             sapList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             basidList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             basidList[['k']] + guides(color=FALSE, shape=FALSE), 
             basidList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             pathList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             pathList[['k']] + guides(color=FALSE, shape=FALSE), 
             pathList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             ooList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             ooList[['k']] + guides(color=FALSE, shape=FALSE), 
             ooList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             ncol=3)

```

*Hyp:* Greater saprotroph and basidiomycete richness will lead to better-fitting decay models (ne.r2), faster decay (k), and less lagginess (alpha) because the community does not need to wait for the arrival of key decayers to act on the wood substrate.  
Hyp-Alt: Greater saprotroph and basidiomycete richness will lead to worse-fitting decay models (ne.r2), slower decay (k), and more lagginess (alpha) because decayers will be allocating more of their resources to combat one another.
**No apparent pattern**

*Hyp:* Greater pathogen and oomycete richness will lead to worse-fitting decay models (ne.r2), slower decay (k), and more lagginess (alpha) because the presence of these organisms will inhibit the establishment and activity of decayers.
**Maybe there's some indicaiton that oomycete presence increases the likelihood of slower (k) and more laggy (alpha) decay**


5. Diversity plus traits as a predictor




##############################################

# Extra pieces

1. *code/testing_time_zero.Rmd* -- Including t=0 points to fit decay model affects the liklihood and the model selection criteria, but the curve fits are identical with this formulation.  Excluding the t=0 fits has an effect of prefering simpler models, which is the same effect as increasing the penalty for model complexity. 

2. *code/initialDist_vs_decayDist_btwCode.Rmd* -- No apparent relationship between species+size dissimilarities in initial microbial community composition (bray and jaccard) and decay trajectory params

3. *code/boralOTUpairs_vs_decay.Rmd* -- No apparent relationship between frequency of boral-ID'd positively/negatively correlated OTU pairs and decay params

4. *code/withinInitialDist_vs_decayR2.Rmd* -- No apparent relationship between initial microbial diversity WITHIN species+size and decay model R2

