---
title: "Community v chemistry -- Recent analyses"
date: "10/23/2017"
output: github_document
---

```{r, include = F}
#chunk options
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

```{r load}
# load libraries and fxns

#libraries
library(tidyverse) # formatting
library(gridExtra) #plotting
#library(litterfitter) #fit decay models to timeseries
#library(GGally) #plotting
#library(corrplot) #plotting
#library(lmtest) # for coxtest()
#library(vegan) #diversity metrics
#library(rioja) #WA-PLS

#fxns
source('code/helper_fxns.R')
sourceDir('code')

```

## Repo outline

### Main analyses

1. How do wood species and sizes vary in decay?
- script: *1_decayPatterns.Rmd*
```{r q1}
#fxns
# make_figs_decayPatterns.R

# load data
stemSamples <- load_stemSamples() # this function is in load_microbeData.R
initial_mass <- read_in_initial_mass()
harvest_mass <- LoadHarvestFiles()
pmr <- Calc_massRemaining(initial_mass, harvest_mass) # calculate percent mass remaining (pmr)
decayfits <- read_csv("derived_data/decayfits.csv") # see rmd script if need to update
# weibull params
# beta = scale param = higher values mean slower decay
# alpha = scale param = higher values mean more S-shaped

# plot potential weibull fits
plot_weibull(pmr, decayfits)

# plot weibull vs neg exp fits
p.a <- makefig__compare_t50_tab(decayfits) 
p.b <- makefig__compare_t50_aic(decayfits) 
grid.arrange(p.a + ggtitle("a"), 
             p.b + ggtitle("b"))

# plot decay fits for each species+size
p.list <- makefig__decayfits(decayfits) 
names(p.list)
makefig__pmr_byStem(pmr, decayfits, param.labels = F) 
makefig__scaleVshape(decayfits)

```

2. Do wood traits explain decay? 
- script: *2_woodtraits_explainDecay.Rmd*
```{r load_woodtraits}
#fxns
# make_figs_decayPatterns.R

# load data
stemSamples <- load_stemSamples() # this function is in load_microbeData.R
initial_mass <- read_in_initial_mass()
harvest_mass <- LoadHarvestFiles()
pmr <- Calc_massRemaining(initial_mass, harvest_mass) # calculate percent mass remaining (pmr)
pmr_byStem <- AvgPMR_byStem(pmr, long.form = F) # average pmr for each stem and timepoint
stem.respVars <- list("time7", "time13", "time25", "time37","time59")
decayfits <- read_csv("derived_data/decayfits.csv") # see 1_decayPatterns.Rmd to update
code.respVars <- list("w.t50","beta","alpha","w.r2")
traits.code <- trait.means_byCode(stemSamples, fill.densitybark = TRUE) # averaged by code; if fill.densitybark == TRUE, then use small stem estimates to approximate large stem density and bark thickness
traits.stem <- trait.means_byStem(stemSamples) # averaged by codeStem
trait.order <- traitcol.order()
traitVars <- c(trait.order$phys.cols, trait.order$nutr.cols)
traitVars.stem <- trait.order$stem.trait.cols
traitVars.cfract <- trait.order$cfract.cols

#######
# do traits predict decayfits?
result.traitsDecay <- doAnalysis_traits_explain_decayParams(decayfits, 
                                      traits.code,
                                      code.respVars, 
                                      traitVars,
                                      cfract = F) # see make_figs_woodTraits_explainDecay.R
fig <- makefig__traits_explain_decayParams(result.list = result.traitsDecay, 
                                           traitVars, cfract = F) 
fig$p
result.traitsDecay$fitstats

#######
# do traits predict PMR?
stem.respVars %>%
  purrr::set_names() %>%
  map(~CreateTraitPMRpair(respVar = .x, 
                          traits.stem, traits.code, 
                          pmr_byStem, 
                          traitVars.stem)) -> datasets # see make_figs_woodTraits_explainDecay.R
result.traitsPMR <- doAnalysis_traits_explain_pmr(datasets, 
                                      stem.respVars, traitVars.stem, 
                                      cfract = F) # see make_figs_woodTraits_explainDecay.R
fig <- makefig__traits_explain_pmr(result.list = result.traitsPMR, 
                                   traitVars.stem,
                                   cfract = F) 
fig$p
result.traitsPMR$fitstats


#######
# do Cfract predict decayfits?
traits.code %>%
  select(c(trait.order$id.cols, trait.order$cfract.cols, "waterperc","N","Ca")) %>%
  filter(!is.na(Arabinose)) -> traits.code.cfract
result.cfractDecay <- doAnalysis_traits_explain_decayParams(decayfits, 
                                      traits.code = traits.code.cfract,
                                      code.respVars, 
                                      traitVars = traitVars.cfract,
                                      cfract = T) # see make_figs_woodTraits_explainDecay.R
fig <- makefig__traits_explain_decayParams(result.list = result.cfractDecay, 
                                           traitVars = traitVars.cfract, 
                                           cfract = T) 
fig$p
result.cfractDecay$fitstats

#######
# do Cfract predict PMR?
traits.stem %>%
  select(c(trait.order$id.cols, "codeStem", trait.order$cfract.cols)) %>%
  filter(!is.na(Arabinose)) -> traits.stem.cfract
stem.respVars %>%
  purrr::set_names() %>%
  map(~CreateCfractPMRpair(respVar = .x, 
                          traits.stem.cfract, 
                          pmr_byStem, 
                          traitVars.cfract)) -> datasets # see make_figs_woodTraits_explainDecay.R
result.cfractPMR <- doAnalysis_traits_explain_pmr(datasets, 
                                      stem.respVars, 
                                      traitVars.stem = traitVars.cfract,
                                      cfract = T) # see make_figs_woodTraits_explainDecay.R
fig <- makefig__traits_explain_pmr(result.list = result.cfractPMR, 
                                   traitVars.stem = traitVars.cfract,
                                   cfract = T) 
fig$p
result.cfractPMR$fitstats

```

3a. Does endophyte composition explain decay?
- script: *3a_endoComp_explainDecay.Rmd*
- summary: No. There is a relationship between composition and percent mass remaining after 37 months, but this is likely driven by the role of water content on the microbial community and decay.
```{r load_microbes}
#fxns
# make_figs_endoComp_explainDecay.R

# load data
stemSamples <- load_stemSamples() # this function is in load_microbeData.R
initial_mass <- read_in_initial_mass()
harvest_mass <- LoadHarvestFiles()
pmr <- Calc_massRemaining(initial_mass, harvest_mass) # calculate percent mass remaining (pmr)
pmr_byStem <- AvgPMR_byStem(pmr, long.form = F) # average pmr for each stem and timepoint
stem.respVars <- list("time7", "time13", "time25", "time37","time59")
decayfits <- read_csv("derived_data/decayfits.csv") # see 1_decayPatterns.Rmd to update
code.respVars <- list("w.t50","beta","alpha","w.r2")
traits.code <- trait.means_byCode(stemSamples, fill.densitybark = TRUE) # averaged by code; if fill.densitybark == TRUE, then use small stem estimates to approximate large stem density and bark thickness
traits.stem <- trait.means_byStem(stemSamples) # averaged by codeStem
trait.order <- traitcol.order()
traitVars <- c(trait.order$phys.cols, trait.order$nutr.cols)
traitVars.stem <- trait.order$stem.trait.cols
traitVars.cfract <- trait.order$cfract.cols
# residuals from 'woodTraits_explainDecay.Rmd'
resids_explainDecay.list <- readRDS(file = "derived_data/resids_explainDecay_list.RData")
# microbes
microb.data <- load_MicrobeCollection(stemSamples)
taxAndFunguild <- microb.data$taxAndFunguild
comm.otu <- microb.data$comm.otu
seqSamples <- microb.data$seqSamples
#check for oomycetes
colnames(comm.otu)[grep("oo", colnames(comm.otu))]
taxAndFunguild %>%
  filter(kingdom != "Fungi")

# identify the saprotrophs
taxAndFunguild %>%
  filter(grepl("Sapro", Guild)) -> sapro.tax
comm.otu.sapro <- comm.otu[, colnames(comm.otu) %in% sapro.tax$OTUId]
dim(comm.otu.sapro) # 176 OTUs
comm.otu.sapro.trimmed <- comm.otu.trimmed[, colnames(comm.otu.trimmed) %in% sapro.tax$OTUId]
dim(comm.otu.sapro.trimmed) # only 11 OTUs

######
# Decay

# 1. decayfits ~ community
cvfit.results.code <- doAnalysis_endoComp_explainDecay(comm.otu, seqSamples, decayfits, code.respVars, use.cache = F, save.cache = F)
cvfit.results.code$prettyTabs #ns
cvfit.results.code.sapro <- doAnalysis_endoComp_explainDecay(comm.otu.sapro, seqSamples, decayfits, code.respVars, use.cache = F, save.cache = F)
cvfit.results.code.sapro$prettyTabs #ns

# 2. pmr ~ community
cvfit.results.stem <- doAnalysis_endoComp_explainPMR(comm.otu, pmr_byStem, stem.respVars)
cvfit.results.stem$prettyTabs # p = 0.04 for time37 full community 
cvfit.results.stem.sapro <- doAnalysis_endoComp_explainPMR(comm.otu.sapro, pmr_byStem, stem.respVars)
cvfit.results.stem.sapro$prettyTabs # ns

######
# Decay trait and Cfract residuals

# Slice 1: Examine Cfract residuals separately

# 1. decayfits trait residuals ~ community
cvfit.results.code <- doAnalysis_endoComp_explainDecayResids(comm.otu, seqSamples, decayfits, code.respVars,traitResiduals.code = resids_explainDecay.list$traits.code, use.cache = F, save.cache = F)
cvfit.results.code$prettyTabs #ns
cvfit.results.code.sapro <- doAnalysis_endoComp_explainDecayResids(comm.otu.sapro, seqSamples, decayfits, code.respVars, traitResiduals.code = resids_explainDecay.list$traits.code, use.cache = F, save.cache = F)
cvfit.results.code.sapro$prettyTabs #ns

# 2. decayfits Cfract residuals ~ community
cvfit.results.code <- doAnalysis_endoComp_explainDecayResids(comm.otu, seqSamples, decayfits, code.respVars, traitResiduals.code = resids_explainDecay.list$cfract.code, use.cache = F, save.cache = F)
cvfit.results.code$prettyTabs #ns
cvfit.results.code.sapro <- doAnalysis_endoComp_explainDecayResids(comm.otu.sapro, seqSamples, decayfits, code.respVars, traitResiduals.code = resids_explainDecay.list$cfract.code, use.cache = F, save.cache = F)
cvfit.results.code.sapro$prettyTabs #ns

# 3. pmr trait residuals ~ community
cvfit.results.stem <- doAnalysis_endoComp_explainPMRResids(comm.otu, pmr_byStem, stem.respVars, traitResiduals.stem = resids_explainDecay.list$traits.stem)
cvfit.results.stem$prettyTabs #ns
cvfit.results.stem.sapro <- doAnalysis_endoComp_explainPMRResids(comm.otu.sapro, pmr_byStem, stem.respVars, traitResiduals.stem = resids_explainDecay.list$traits.stem)
cvfit.results.stem.sapro$prettyTabs #ns


# Slice 2: Include C fraction data (and exclude small stems)

# 1. decayfits trait residuals ~ community
cvfit.results.code <- doAnalysis_endoComp_explainDecayResids(comm.otu, seqSamples, decayfits, code.respVars, traitResiduals.code = resids_explainDecay.list$cfract.code, use.cache = F, save.cache = F)
cvfit.results.code$prettyTabs # ns
cvfit.results.code.sapro <- doAnalysis_endoComp_explainDecayResids(comm.otu.sapro, seqSamples, decayfits, code.respVars, traitResiduals.code = resids_explainDecay.list$cfract.code, use.cache = F, save.cache = F)
cvfit.results.code.sapro$prettyTabs # ns

# 2. pmr Cfract residuals ~ community
cvfit.results.stem <- doAnalysis_endoComp_explainPMRResids(comm.otu, pmr_byStem, stem.respVars, traitResiduals.stem = resids_explainDecay.list$cfract.stem)
cvfit.results.stem$prettyTabs # ns
cvfit.results.stem.sapro <- doAnalysis_endoComp_explainPMRResids(comm.otu.sapro, pmr_byStem, stem.respVars, traitResiduals.stem = resids_explainDecay.list$cfract.stem)
cvfit.results.stem.sapro$prettyTabs # p=0.06 for time59 full saprotroph community

```

3b. Does endophyte diversity explain decay? *
- script: *3b_endoDiv_explainDecay.Rmd*
*** revisit this - add bonferonni correction for multiple comparisons ***

```{r load_microbesdiv}


```



### Other pieces

- How does including t=0 points affect the decay model fit? Including it affects the liklihood and the model selection criteria, but the curve fits are identical.  Excluding t=0 leads to prefering simpler models, which is the same effect as increasing the penalty for model complexity -- *code/testing_time_zero.Rmd* 

- Do endophyte community dissimilarities correlate with decay trajectory parameters? No relationships -- *code/initialDist_vs_decayDist_btwCode.Rmd*

- Is there a relationship between the frequency of highly correlated (negative or positive) OTU pairs identified via boral and decay parameters? Nope -- *code/boralOTUpairs_vs_decay.Rmd*

- Is there a relationship between initial microbial diversity WITHIN species+size categories and decay model R2? Nope -- *code/withinInitialDist_vs_decayR2.Rmd*

- Exclude mycorrhizal fungi and animal-associated fungi that somehow made it into our OTU table -- *code/unexpectedTaxa.Rmd*


### Need to continue incorporating C fract analyses into "woodTraits_explainDecay.Rmd"

Check out correlations among C fractions and other traits. Exclude two samples for which there isn't waterperc data. Remember that there is no stem-level barkthick or density data for large stem samples.
```{r}
cfract.load<- load_Cfract()
cfract.load %>%
  select(-Total) -> cfract.load
traits.stem %>%
  left_join(cfract.load) %>%
  filter(!is.na(Arabinose)) %>%
  select(-c(barkthick, density)) -> traits.stem.sel
#two samples are missing waterperc data
traits.stem.sel %>%
  filter(is.na(waterperc)) %>%
  select(codeStem, waterperc) 
#remove them
traits.stem.sel %>%
  filter(!is.na(waterperc)) -> traits.stem.sel

#grab just the matrix of traits
traits.stem.sel %>%
  select(-c(species, size, code, unique, Stem)) -> matonly
matonly.s <- scale(as.matrix(matonly[,-1]))

# plot correlations
pdf(file = "output/corplot.pdf")
corrplot(cor(matonly.s), type = "lower")
dev.off()

```

Kick out traits that are highly correlated
- C and perc.Total.lignin are highly positive correlated (0.81); keep perc.Total.lignin
- Glucose+Xylose and perc.Total.lignin are highly negatively correlated (-.89, -.73); keep perc.Total.lignin.
```{r}
#update
traits.stem.sel %>%
  select(-c(Glucose, Xylose, C)) -> traits.stem.sel
traits.stem.sel %>%
  select(-c(species, size, code, unique, Stem)) -> matonly
matonly.s <- scale(as.matrix(matonly[,-1]))

# plot correlations again
corrplot(cor(matonly.s), type = "lower")

cor(matonly.s)

```

Plot variation in C fractions w/in and between species
```{r}

traits.stem.sel %>%
  select(-c(Ca, Fe, K, Mn, N, P, waterperc, Zn)) %>%
  gather(key = "trait", value = "trait.val", -c(codeStem, species, size, code, unique, Stem)) -> tmp

p <- ggplot(tmp, aes(x = reorder(species, trait.val), y = trait.val)) +
  geom_point() +
  facet_wrap(~trait, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave(p, filename = "output/cfract_univar.pdf")

```

Check out samples in trait space
```{r}
pca <- prcomp(matonly.s)
df <- data.frame(pca$x[,1:2], 
                 traits.stem.sel[,c("codeStem","species","code","size")])
datapc <- data.frame(varnames=rownames(pca$rotation), pca$rotation[,1:2])
mult <- min(
  (max(df[,"PC1"]) - min(df[,"PC2"])/(max(datapc[,"PC2"])-min(datapc[,"PC2"]))),
  (max(df[,"PC2"]) - min(df[,"PC1"])/(max(datapc[,"PC1"])-min(datapc[,"PC1"])))
)
datapc <- transform(datapc,
                    v1 = .7 * mult * datapc$PC1,
                    v2 = .7 * mult * datapc$PC2)
  

plot_12 <- ggplot(data=datapc, aes(x=v1, y=v2, label=varnames)) + 
  geom_text(size=3, vjust=1, color=1) + 
  geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2),
               arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color=1) + 
  coord_fixed() +
  geom_text(mapping=aes(x=PC1, y=PC2, 
                        color = species, label = code), data=df, inherit.aes = FALSE) +
  theme_bw() + xlab("PC1 (24%)") + ylab("PC2 (16%)") +
  guides(color = F)
plot_12
ggsave(plot_12, filename = "output/stem_traitord.pdf")
# for some reason this plot knit with more extra data points? it seems to work fine in ggsave

summary(pca)
```
![](output/stem_traitord.pdf)

Load new versions of trait data
```{r}
traits.code <- trait.means_byCode_new(stemSamples, fill.densitybark = T)
traits.stem <- trait.means_byStem_new(stemSamples)

# remove traits that are highly correlated with others
exclude <- c("Glucose","Xylose","C")
traits.code %>%
  select(-exclude) -> traits.code
traits.stem %>%
  select(-exclude) -> traits.stem

# remove observations w/o C fract data
traits.code %>%
  filter(!is.na(Arabinose)) -> traits.code.tmp
traits.stem %>%
  filter(!is.na(Arabinose)) -> traits.stem.tmp

# remove density and barkthickness from stem data
traits.stem.tmp %>%
  select(-c("barkthick","density")) -> traits.stem.tmp

```

Do wood C fractions explain decay?

(by code)
Note that there are only 12 observations by code, so that limits the number of traits we can include in the model. Not including micronutrients

```{r}
#merge traits and response variables into 1 df
decayfits %>%
  select(code, species, size, 
         ne.r2, k, t50, 
         w.r2, alpha, beta, w.t50) %>%
  left_join(traits.code) %>% 
  filter(!is.na(P)) %>%
  filter(!is.na(waterperc)) -> decayfits.traits
# subset by wood fract data
decayfits.traits %>%
  filter(!is.na(Arabinose)) -> decayfits.traits

# scale all the traits
decayfits.traits %>%
  select(waterperc, N, P, perc.Total.lignin, Rhamnose, Arabinose, Galactose, Mannose) -> mat
decayfits.traits %>%
  select(-c(waterperc, N, P, perc.Total.lignin, Rhamnose, Arabinose, Galactose, Mannose)) -> tmp
decayfits.traits.s <- data.frame(tmp, scale(mat))

#set up full models
respVars <- list("k","t50","ne.r2", "alpha","beta","w.t50","w.r2")
rhs <- "waterperc + N + P + perc.Total.lignin + Rhamnose + Arabinose + Galactose + Mannose" # got rid of size because only looking at large stems
mod.full.list<-lapply(respVars, ModelFit_manyYs, rhs, curr.data=decayfits.traits.s) # summaryTable_fxns.R

#do stepwise model selection #uncomment if data changes
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="backward")
  return(mod.select)
  })
names(mod.select.list) <- respVars

#plot
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(sum.list, unlist(respVars))
fitstat.df<-PullLmFitStats(sum.list, unlist(respVars))
fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
coefs.df %>%
  left_join(r2.df) %>%
  filter(respvar %in% c("alpha","beta","w.t50","w.r2")) -> coefs.df

unique(coefs.df$term)
# term.order <- c("(Intercept)","waterperc","N","perc.Total.lignin","Arabinose","Galactose","Mannose","Rhamnose")
# coefs.df$term <- factor(coefs.df$term, levels = rev(term.order))

coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_wrap(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p
ggsave(p, filename = "output/cfract_code.pdf")

```


(by stem) -- I want to update this so that the traits values are scaled before running the models
```{r}
# create dataframes
# stem-level waterperc, xrf, and CN and (small) species-level barkthickness and density

respVars <- list("time7", "time13", "time25", "time37","time59")

traits.stem.tmp %>%
  filter(!is.na(P)) %>%
  filter(!is.na(waterperc)) %>%
  filter(!is.na(Arabinose)) -> traits.stem.tmp

# scale all the traits
traits.stem.tmp %>%
  select(waterperc, N, P, perc.Total.lignin, Rhamnose, Arabinose, Galactose, Mannose) -> mat
traits.stem.tmp %>%
  select(codeStem) -> tmp
traits.stem.tmp.s <- data.frame(tmp, scale(mat[,-1]))
traits.stem.tmp.s

datasets<-lapply(respVars, function(x) {
  result<-CreateTraitPMRpair(x, traits.stem, traits.code, pmr_byStem) #analysisDF_fxns.R
  return(result)
})
names(datasets)<-respVars

#set up full models
rhs <- "waterperc + N + Arabinose + Galactose + Mannose + perc.Total.lignin + Rhamnose"
lhs <- "curr.pmr"
mod.full.list<-lapply(datasets, function(x) {
  result<-ModelFit_manyYs(y=lhs, rhs=rhs, curr.data=x)
  return(result)
})

#do stepwise model selection
mod.select.list<-lapply(mod.full.list, function(x) {
  x.updated<-update(x, . ~ ., data = model.frame(x))
  mod.select<-step(x.updated, direction="backward")
  return(mod.select)
  })

#plot
sum.list<-lapply(mod.select.list, summary)
coefs.df<-PullLmCoefs(sum.list, unlist(respVars))
fitstat.df<-PullLmFitStats(sum.list, unlist(respVars))
fitstat.df %>%
  gather(key = "respvar", value = "value", -c(term)) %>%
  filter(term == "r.squared") %>%
  select(-term) %>%
  rename('r.squared'=value) -> r2.df
coefs.df %>%
  left_join(r2.df) -> coefs.df

coefs.df$respvar<- factor(coefs.df$respvar, levels = c("time7","time13","time25","time37","time59"))
coefs.df
term.order <- c("(Intercept)","waterperc","N","perc.Total.lignin","Arabinose","Galactose","Mannose","Rhamnose")
coefs.df$term <- factor(coefs.df$term, levels = rev(term.order))

coefs.df %>%
  filter(term != "(Intercept)") -> tmp
p.stem <- ggplot(tmp, aes(x = est, y = term, color = stars)) +
  geom_point() +
  geom_errorbarh(aes(xmin = est - se, xmax = est + se)) +
  facet_grid(~respvar, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_color_manual(values = c("gray","black","blue","red")) +
  ylab("Predictor variable") +
  xlab("Coef. estimate (+/- SE)")
p.stem

ggsave(p.stem, filename = "output/cfract_stem.pdf")




```


Visualize the relationship between wood C fractions and endophyte communities
```{r}

#microbes
microb.data <- load_MicrobeCollection(stemSamples)
taxAndFunguild <- microb.data$taxAndFunguild
comm.otu <- microb.data$comm.otu
seqSamples <- microb.data$seqSamples

cfract.load<- load_Cfract()
mat.otu <- sub.comm.otu <- comm.otu[row.names(comm.otu) %in% cfract.load$unique,]
dim(mat.otu)

xVars <- cfract.load
xVars %>%
  select(-c(Total, unique, Stem, size)) -> envVars
dim(envVars)  

# transform otu tab
mat.otu.t<-decostand(mat.otu, 'hellinger')
  
# unconstrained mod
mod.obj <- capscale(mat.otu.t ~ 1, data = envVars, distance='bray')
  
  # do envfit
  envfit.obj <- envfit(mod.obj, envVars)
  
  # collect site scores from mod.obj
  scrs <- as.data.frame(scores(mod.obj, display = "sites"))
  scrs
  scrs <- cbind(scrs, seq_sampName = row.names(scrs))
  scrs %>%
    left_join(seqSamples) -> scrs
  
  # collect vector scores from envfit.obj
  spp.scrs <- as.data.frame(scores(envfit.obj, display = "vectors"))
  spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), pval = envfit.obj$vectors$pvals)
  
  # plot
  mult <- 2 #multiplier for the arrows and text for envfit
  mult.text <- 2.3
  p <- ggplot(scrs) +
    geom_text(mapping = aes(x = MDS1, y = MDS2, label = code,
                             color = species), size = 3) +
    coord_fixed() + ## need aspect ratio of 1!
    geom_segment(data = spp.scrs,
                 aes(x = 0, xend = mult*MDS1, y = 0, yend = mult*MDS2),
                 arrow = arrow(length = unit(0.2, "cm")), colour = 1) +
    geom_text(data = spp.scrs, 
              aes(x = mult.text*MDS1, y = mult.text*MDS2, 
                  label = Species), size = 4, fontface = "bold") +
    xlab("PCoA 1") + ylab("PCoA 2")
  p
  ggsave(file="output/unconstrained_envfit.pdf", width = 8, height = 5)

```

Does adding C fraction data explain more variation in endophyte composition? How much? Doubles the unique info (with the caveat that I didn't include micronutrients)

X1 = original trait set: waterperc, C, N, and P
X2 = C fractions: perc.Total.lignin, Rhamnose, Arabinose, Galactose, Mannose
```{r}

traits.stem <- trait.means_byStem_new(stemSamples)

# remove traits that are highly correlated with others
exclude <- c("Glucose","Xylose") # don't get rid of C
traits.stem %>%
  select(-exclude) -> traits.stem

# remove observations w/o C fract data
traits.stem %>%
  filter(!is.na(Arabinose)) -> traits.stem.tmp

# remove density and barkthickness from stem data
traits.stem.tmp %>%
  select(-c("barkthick","density")) -> traits.stem.tmp

traits.stem.tmp %>%
  filter(!is.na(P)) %>%
  filter(!is.na(waterperc)) %>%
  filter(!is.na(Arabinose)) -> traits.stem.tmp

# scale all the traits
traits.stem.tmp %>%
  select(waterperc, N, P,C, perc.Total.lignin, Rhamnose, Arabinose, Galactose, Mannose) -> mat
traits.stem.tmp %>%
  select(codeStem) -> tmp
traits.stem.tmp.s <- data.frame(tmp, scale(mat[,-1]))
traits.stem.tmp.s

colnames(traits.stem.tmp.s)
envVars <- traits.stem.tmp.s

# remove communities with missing waterperc data
mat.otu <- sub.comm.otu <- comm.otu[row.names(comm.otu) %in% envVars$codeStem,]
dim(mat.otu)
dim(envVars)

# transform otu tab
mat.otu.t<-decostand(mat.otu, 'hellinger')

# extract the unconstrained and constrained mod
 # unconstrained mod

mod0 <- capscale(mat.otu.t ~ 1, data = envVars[,-1], distance='bray')
  # constrained mod
modfull <- capscale(mat.otu.t ~ ., data=envVars[,-1], distance='bray')

# model selection
modstep <- ordistep(mod0, scope=formula(modfull))
  
# summary table for the best model
df <- data.frame(term=row.names(modstep$anova), modstep$anova)
df %>%
    separate(term, into=c("drop","term")) %>%
    select(-drop) %>%
    rename('pval'=`Pr..F.`) -> df
df
write.csv(df, file = "output/constrained_bestmodelsummaryAllTraits.csv")

envVars %>%
  select(waterperc, N, P,C) -> orig.traits
envVars %>%
  select(perc.Total.lignin, Rhamnose, Arabinose, Galactose, Mannose) -> C.fract

endo.var <- varpart(vegdist(mat.otu.t, distance = 'bray'),
                        ~ as.matrix(orig.traits),
                        ~ as.matrix(C.fract))
    var.fract <- endo.var$part
    plot(endo.var)
    
    orig.traits
    C.fract





```



