---
title: "Does chemistry or community better predict mass loss?"
author: "Marissa Lee"
date: "10/23/2017"
output: github_document
---

```{r,message=FALSE}
#chunk options
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

#libraries
devtools::install_github("cornwell-lab-unsw/litterfitter")
library(dplyr)
library(ggplot2)
library(readr)
library(vegan)
library(knitr)
library(litterfitter)
library(magrittr)
library(tidyr)
library(gridExtra)
library(rioja)

#fxns
source("code/load_fxns.R")
source("code/curve_fitting_fxns.R")
source("code/distance_fxns.R")
source("code/otuIDs_fxns.R")


### LOAD MICROBIAL COMMUNITY DATA

#stem sample meta data
#stemSamples<-load_stemSamples() #uncomment if the data changes
#write_csv(stemSamples, "derived_data/stemSamples.csv")
stemSamples<-read_csv("derived_data/stemSamples.csv")

#OTU table
#fung.otu<-load_matotu() #uncomment if the data changes
#comm.otu<-add_oomycetes(fung.otu) #add the oomycetes #uncomment if the data changes
#write.csv(comm.otu, "derived_data/comm_otu.csv")
comm.otu<-read.csv("derived_data/comm_otu.csv", row.names=1)

#create sequence sample meta data table
#seqSamples<-load_seqSamples(comm.otu, stemSamples) #uncomment if the data changes
#write_csv(seqSamples, "derived_data/seqSamples.csv")
seqSamples<-read_csv("derived_data/seqSamples.csv")

#taxon lookup info
#taxAndFunguild<-load_TaxAndFunguild(comm.otu) #uncomment if the data changes
#write_csv(taxAndFunguild, "derived_data/taxaAndFunguild.csv")
taxAndFunguild<-read_csv("derived_data/taxaAndFunguild.csv")

#plot_sampleEffortCurves(comm.otu)


### LOAD WOOD TRAIT DATA
#traits.mean<-mergeTraitData() #uncomment if the data changes
#write_csv(traits.mean, "derived_data/traits_mean.csv") 
traits.mean<-read_csv("derived_data/traits_mean.csv")
#missing data
#traits.long<-as.data.frame(gather(traits.mean, key=trait, value=value, -(1:3)))
#filter(traits.long, is.na(value))


### LOAD MASS LOSS DATA and CALCULATE % MASS REMAINING AT EACH TIMEPOINT
#initial_mass <- read_in_initial_mass() #uncomment if the data changes
#harvest_mass<-LoadHarvestFiles()
#mass.data<-bind_rows(initial_mass, harvest_mass)
#missing data
#mass.data %>% filter(is.na(totalSampleDryMass))
#plotting_df<-Calc_massRemaining(mass.data)
#matching failures
# plotting_df %>%
#   filter(is.na(pmr)) %>%
#   select(unique, species, size, time, totalSampleDryMass, notes) %>%
#   spread(key=time, value=totalSampleDryMass)
#remove NAs
#plotting_df %>% filter(!is.na(pmr)) -> plotting_df
#write_csv(plotting_df,"derived_data/plotting_df.csv")
plotting_df<-read_csv("derived_data/plotting_df.csv")

### CALCULATE DECAY TRAJECTORY FITS
#spdf <- fit_all_curves(plotting_df) #this recalculates all the curve fits, uncomment if the data changes
#indx<-select(stemSamples, code, species, size)
#spdf<-left_join(spdf, indx) #add code
#write_csv(spdf,"derived_data/mass_loss_parameters.csv")
spdf <- read_csv("derived_data/mass_loss_parameters.csv")
# ggplot(spdf,aes(x=t70,y=w.t70,col=size))+
#   geom_point()+
#   labs(x="Time to 30% mass loss (negative exponential)", 
#        y="Time to 30% mass loss (Weibull)")+
#   geom_abline(slope=1,intercept=0,linetype="dashed")+theme_bw()

```



1. Wood traits as a preditor
*Hyp:* Variation in wood traits will lead to differences in decay model fit (r2), rate (k), and lagginess (alpha).  Specifically, we expect samples with (a) high waterperc, (b) low density and C, (c) high P, K, Ca, Mn, Fe, Zn, and N, and (d) thicker bark (potential mech: limiting microbial colonization) to have better-fiting decay models (r2), faster decay rates (k), and less lagginess (alpha). 
```{r, include=FALSE}

#merge into 1 df
spdf %>%
  select(code, species, size, ne.r2, k, t70, alpha) %>%
  left_join(traits.mean) %>% 
  filter(!is.na(P)) %>%
  filter(!is.na(waterperc)) -> spdf.traits

#fit full models
mod.full.r<-lm(ne.r2 ~ size + waterperc + density + barkthick + P + K + Ca + Mn + Fe + Zn + N + C, data=spdf.traits)
mod.full.k<-lm(k ~ size + waterperc + density + barkthick + P + K + Ca + Mn + Fe + Zn + N + C, data=spdf.traits)
mod.full.t70<-lm(t70 ~ size + waterperc + density + barkthick + P + K + Ca + Mn + Fe + Zn + N + C, data=spdf.traits)
mod.full.alpha<-lm(alpha ~ size + waterperc + density + barkthick + P + K + Ca + Mn + Fe + Zn + N + C, data=spdf.traits)

#do stepwise model selection
mod.select.r<-step(mod.full.r, direction="backward")
mod.select.k<-step(mod.full.k, direction="backward")
mod.select.t70<-step(mod.full.t70, direction="backward")
mod.select.alpha<-step(mod.full.alpha, direction="backward")

#save the residuals
trait.residuals<-data.frame(code=spdf.traits$code,
  r2.resid=mod.select.r$residuals,
  k.resid=mod.select.k$residuals,
  t70.resid=mod.select.t70$residuals,
  alpha.resid=mod.select.alpha$residuals)

```

*Result:*
- r2... greater water content and greater Zn and N leads to better-fitting decay models. Note: This result changed when I changed waterperc to g water/g wet weight. When waterperc was in terms of g/g dry weight, the best model that greater water content and less C leads to better-fitting decay models
```{r}
summary(mod.select.r) # waterperc, Zn, N
#ggplot(spdf.traits, aes(x=waterperc, y=ne.r2, color=species, shape=size)) + geom_point()
```
- k... small size stems, greater water content, thinner bark, less Ca, more Zn, and more N lead to faster decay
```{r}
summary(mod.select.k) # size, waterperc, barkthick, Ca, Zn, N
ggplot(spdf.traits, aes(x=waterperc, y=k, color=species, size=density)) + geom_point() + facet_grid(~size) +
  xlab("Water content (% wet weight)")
#ggplot(spdf.traits, aes(x=waterperc, y=density, color=species, size=k)) + geom_point() + facet_grid(~size)
```
- t70... small stem sizes, less water content, thicker bark, more Ca, less Zn, and less N lead to longer wood "70%"-lives.
Note: This result changed when I changed waterperc to g water/g wet weight. When waterperc was in terms of g/g dry weight, the best model indicated that large size stems, less water content, more Ca, and less Zn lead to longer wood "70%"-lives
```{r}
summary(mod.select.t70) # size, waterperc, barkthick, Ca, Zn
#ggplot(spdf.traits, aes(x=waterperc, y=t70, color=species, size=density)) + geom_point() + facet_grid(~size)
```
- alpha--- note: don't interpret yet
```{r}
summary(mod.select.alpha) # density, Zn, C
```




2. Community as a predictor
*Hyp:* Average initial microbial communitiy compositions will predict variation in decay model fit (r2), rate (k), and lagginess (alpha).  
```{r, include=FALSE}

#average OTU abundances by code
CODE<-unique(seqSamples$code)
meanOTUabund.list<-list()
sdOTUabund.list<-list()
for(i in 1:length(CODE)){
  #identify the seq_sampNames to aggregate
  curr.seqSamps<-seqSamples[seqSamples$code==CODE[i],]$seq_sampName 
  #filter 
  curr.comm.otu<-comm.otu[row.names(comm.otu) %in% curr.seqSamps,]
  #summarize
  meanOTUabund.list[[i]]<-apply(curr.comm.otu, 2, mean, na.rm=TRUE)
  sdOTUabund.list[[i]]<-apply(curr.comm.otu, 2, sd, na.rm=TRUE)
}
names(meanOTUabund.list)<-CODE
names(sdOTUabund.list)<-CODE
meanOTUabund.list %>% combine() %>% as.data.frame %>% t() -> 
  meanOTUabund
sdOTUabund.list %>% combine() %>% as.data.frame %>% t() -> 
  sdOTUabund

#get rid of empty OTU cols
#sum(colSums(meanOTUabund)!=0)
meanOTUabund.trim<-meanOTUabund[,colSums(meanOTUabund)!=0]

#get rid of spdf rows for which there is missing community data... this is eusc
#spdf[!spdf$code %in% row.names(meanOTUabund.trim),]$code
spdf.trim<-spdf[spdf$code %in% row.names(meanOTUabund.trim),]

#fit models
fit.r2 <- WAPLS(meanOTUabund.trim, spdf.trim$ne.r2)
fit.k <- WAPLS(meanOTUabund.trim, spdf.trim$k)
fit.t70 <- WAPLS(meanOTUabund.trim, spdf.trim$t70)
fit.alpha <- WAPLS(meanOTUabund.trim, spdf.trim$alpha)

#cross-validate models using the leave-one-out method
fit.r2.cv <- crossval(fit.r2, cv.method="loo")
fit.k.cv <- crossval(fit.k, cv.method="loo")
fit.t70.cv <- crossval(fit.t70, cv.method="loo")
fit.alpha.cv <- crossval(fit.alpha, cv.method="loo")

#rand.t.test(fit.alpha.cv) #perform randomization t-test to test the significance of a cross-validated model
#screeplot(fit.r2.cv)

```
*Result:*
-r2... none of the community components are significant predictors
```{r}
rand.t.test(fit.r2.cv)
```
-k... none of the community components are significant predictors
```{r}
rand.t.test(fit.k.cv)
```
-t70... none of the community components are significant predictors
```{r}
rand.t.test(fit.t70.cv)
```
-alpha --- note: don't interpret yet
```{r}
rand.t.test(fit.alpha.cv)
```


3. Community+traits as a predictor
```{r}
#run rioja::WAPLS on wood trait residuals

#make sure that the residuals and community dataframes are aligned
trait.residuals$code<-as.character(trait.residuals$code)
trait.residuals[!trait.residuals$code %in% row.names(meanOTUabund.trim),"code"] # no missing from meanOTUabund.trim
row.names(meanOTUabund.trim)[!row.names(meanOTUabund.trim) %in% trait.residuals$code] #missing olst from trait.residuals
trait.residuals.trim<-trait.residuals[trait.residuals$code %in% row.names(meanOTUabund.trim),] #trim
meanOTUabund.trim2<-meanOTUabund.trim[row.names(meanOTUabund.trim) %in% trait.residuals.trim$code,] #trim
ord<-match(row.names(meanOTUabund.trim2), trait.residuals.trim$code)
trait.residuals.trim.o<-trait.residuals.trim[ord,]
sum(trait.residuals.trim.o$code != row.names(meanOTUabund.trim2)) #this needs to by 0

#get rid of empty OTU cols
sum(colSums(meanOTUabund.trim2)!=0)
meanOTUabund.trim2<-meanOTUabund.trim2[,colSums(meanOTUabund.trim2)!=0]

#fit models
fit.tr.r2 <- WAPLS(meanOTUabund.trim2, trait.residuals.trim.o$r2.resid)
fit.tr.k <- WAPLS(meanOTUabund.trim2, trait.residuals.trim.o$k.resid)
fit.tr.t70 <- WAPLS(meanOTUabund.trim2, trait.residuals.trim.o$t70.resid)
fit.tr.alpha <- WAPLS(meanOTUabund.trim2, trait.residuals.trim.o$alpha.resid)

#cross-validate models using the leave-one-out method
fit.tr.r2.cv <- crossval(fit.tr.r2, cv.method="loo")
fit.tr.k.cv <- crossval(fit.tr.k, cv.method="loo")
fit.tr.t70.cv <- crossval(fit.tr.t70, cv.method="loo")
fit.tr.alpha.cv <- crossval(fit.tr.alpha, cv.method="loo")

#rand.t.test(fit.alpha.cv) #perform randomization t-test to test the significance of a cross-validated model
#screeplot(fit.r2.cv)

```
*Result:*
-r2...none of the community components are significant predictors. Note: This result changed when I changed waterperc to g water/g wet weight. When waterperc was in terms of g/g dry weight, Comp05 was marginally significant
```{r}
rand.t.test(fit.tr.r2.cv)
```
-k... none of the community components are significant predictors
```{r}
rand.t.test(fit.tr.k.cv)
```
-t70... none of the community components are significant predictors. Note: This result changed when I changed waterperc to g water/g wet weight. When waterperc was in terms of g/g dry weight, Comp05 was a significant predictor
```{r}
rand.t.test(fit.tr.t70.cv) 
```
-alpha --- note: don't interpret yet
```{r}
rand.t.test(fit.tr.alpha.cv)
```

*Note: no longer warrented based on analyses with waterperc represented as g water / g wet weight...*
Investigate the biology underlying t70-associated coefs for Comp05
```{r}
coef.comp5<-fit.tr.t70.cv$coefficients[,'Comp05']
coef.comp5.df<-data.frame(OTUId=names(coef.comp5), coefComp5=coef.comp5)

#create df of OTU taxon and guild info matched with coef value
coef.comp5.df %>%
  left_join(taxAndFunguild) -> coef.comp5.ann
```
By trophic mode (note that apparently empty cateogies have at least 1 data point, the violin plot just doesn't show it)
```{r}
ggplot(coef.comp5.ann, aes(x=Trophic.Mode, y=coefComp5)) + geom_violin() + coord_flip()
```
By guild (note that apparently empty cateogies have at least 1 data point, the violin plot just doesn't show it)
```{r}
ggplot(coef.comp5.ann, aes(x=Guild, y=coefComp5)) + geom_violin() + coord_flip()
```
By phylum
```{r}
ggplot(coef.comp5.ann, aes(x=phylum, y=coefComp5)) + geom_violin() + coord_flip()
```

4. Diversity (and diversity of specific clades) as a predictor

*Hyp:* Greater microbial diversity (richness, Shannon diversity, phylogenetic diversity) will lead to better-fitting decay models (ne.r2), faster decay (k), and less lagginess (alpha) because of the selection effect for fast decayers and complementarity among taxa for decay.  
Hyp-Alt: Greater microbial diversity will lead to worse-fitting decay models (ne.r2), slower decay (k), and more lagginess (alpha) because taxa will be allocating more of their resources to combat one another.
```{r}


```


*Hyp:* Greater saprotroph and basidiomycete richness will lead to better-fitting decay models (ne.r2), faster decay (k), and less lagginess (alpha) because the community does not need to wait for the arrival of key decayers to act on the wood substrate.  
Hyp-Alt: Greater saprotroph and basidiomycete richness will lead to worse-fitting decay models (ne.r2), slower decay (k), and more lagginess (alpha) because decayers will be allocating more of their resources to combat one another.
**No significant relationships**
```{r}

# summarize the presence of ... in each sample
sapro.df<-Calc_richOTUtype(colNam="Trophic.Mode", grepTerm="Sapro", taxAndFunguild, comm.otu)  
basidio.df<-Calc_richOTUtype(colNam="phylum", grepTerm="Basid", taxAndFunguild, comm.otu)

# create a merged df wtih spdf
saprorich.spdf<-Create_rich_spdf_DF(otutype.df=sapro.df, spdf)
basidrich.spdf<-Create_rich_spdf_DF(otutype.df=basidio.df, spdf)

# fit models
mod.sapro.r2<-lm(ne.r2~size+mean, data=saprorich.spdf)
mod.sapro.k<-lm(k~size+mean, data=saprorich.spdf)
mod.sapro.t70<-lm(t70~size+mean, data=saprorich.spdf)
mod.sapro.alpha<-lm(alpha~size+mean, data=saprorich.spdf)
mod.basid.r2<-lm(ne.r2~size+mean, data=basidrich.spdf)
mod.basid.k<-lm(k~size+mean, data=basidrich.spdf)
mod.basid.t70<-lm(t70~size+mean, data=basidrich.spdf)
mod.basid.alpha<-lm(alpha~size+mean, data=basidrich.spdf)

# evaluate models
summary(mod.sapro.r2)
summary(mod.sapro.k)
summary(mod.sapro.t70)
summary(mod.sapro.alpha)
summary(mod.basid.r2)
summary(mod.basid.k)
summary(mod.basid.t70)
summary(mod.basid.alpha)

# create plots
sapList<-Plot_richOTUtype(rich.spdf=saprorich.spdf, 
                        valueCol_vec=c("ne.r2", "k","alpha"), 
                        otutypeNam="Saprotroph")
basidList<-Plot_richOTUtype(rich.spdf = basidrich.spdf, 
                        valueCol_vec=c("ne.r2", "k","alpha"), 
                        otutypeNam="Basidio")

#plot
grid.arrange(sapList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             sapList[['k']] + guides(color=FALSE, shape=FALSE), 
             sapList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             basidList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             basidList[['k']] + guides(color=FALSE, shape=FALSE), 
             basidList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             ncol=3)

```


*Hyp:* Greater pathogen and oomycete richness will lead to worse-fitting decay models (ne.r2), slower decay (k), and more lagginess (alpha) because the presence of these organisms will inhibit the establishment and activity of decayers.
**Maybe there's some indicaiton that oomycete presence increases the likelihood of slower (k) and more laggy (alpha) decay**
```{r}

# summarize the presence of ... in each sample
path.df<-Calc_richOTUtype(colNam="Trophic.Mode", grepTerm="Patho", taxAndFunguild, comm.otu)
oomy.df<-Calc_richOTUtype(colNam="kingdom", grepTerm="Protist", taxAndFunguild, comm.otu)

# create a merged df wtih spdf
pathrich.spdf<-Create_rich_spdf_DF(otutype.df=path.df, spdf)
oomyrich.spdf<-Create_rich_spdf_DF(otutype.df=oomy.df, spdf)

# # fit models
# mod.sapro.r2<-lm(ne.r2~size+mean, data=saprorich.spdf)
# mod.sapro.k<-lm(k~size+mean, data=saprorich.spdf)
# mod.sapro.t70<-lm(t70~size+mean, data=saprorich.spdf)
# mod.sapro.alpha<-lm(alpha~size+mean, data=saprorich.spdf)
# mod.basid.r2<-lm(ne.r2~size+mean, data=basidrich.spdf)
# mod.basid.k<-lm(k~size+mean, data=basidrich.spdf)
# mod.basid.t70<-lm(t70~size+mean, data=basidrich.spdf)
# mod.basid.alpha<-lm(alpha~size+mean, data=basidrich.spdf)
# 
# # evaluate models
# summary(mod.sapro.r2)
# summary(mod.sapro.k)
# summary(mod.sapro.t70)
# summary(mod.sapro.alpha)
# summary(mod.basid.r2)
# summary(mod.basid.k)
# summary(mod.basid.t70)
# summary(mod.basid.alpha)

# create plots
pathList<-Plot_richOTUtype(rich.spdf=pathrich.spdf, 
                        valueCol_vec=c("ne.r2", "k","alpha"), 
                        otutypeNam="Pathogen")
ooList<-Plot_richOTUtype(rich.spdf=oomyrich.spdf, 
                        valueCol_vec=c("ne.r2", "k","alpha"), 
                        otutypeNam="Oomycete")

# plot
grid.arrange(pathList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             pathList[['k']] + guides(color=FALSE, shape=FALSE), 
             pathList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             ooList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             ooList[['k']] + guides(color=FALSE, shape=FALSE), 
             ooList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             ncol=3)

```

5. Diversity plus traits as a predictor




##############################################

# Extra pieces

1. *code/testing_time_zero.Rmd* -- Including t=0 points to fit decay model affects the liklihood and the model selection criteria, but the curve fits are identical with this formulation.  Excluding the t=0 fits has an effect of prefering simpler models, which is the same effect as increasing the penalty for model complexity. 

2. *code/initialDist_vs_decayDist_btwCode.Rmd* -- No apparent relationship between species+size dissimilarities in initial microbial community composition (bray and jaccard) and decay trajectory params

3. *code/boralOTUpairs_vs_decay.Rmd* -- No apparent relationship between frequency of boral-ID'd positively/negatively correlated OTU pairs and decay params

4. *code/withinInitialDist_vs_decayR2.Rmd* -- No apparent relationship between initial microbial diversity WITHIN species+size and decay model R2

