---
title: "Does the initial frequency of significantly positively/negatively correlated boral-identified OTU pairs explain decay? Nope."
date: "11/7/2017"
output: github_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)



spdf <- read_csv("../derived_data/mass_loss_parameters.csv")
seqSamples<-read_csv("../derived_data/seqSamples.csv")

```

Hyp: A higher frequency of postively-correlated OTUs (after controlling for wood trait environmental influences) in a sample community will lead to worse-fitting decay models (ne.r2), slower decay (k), and more lagginess (alpha).  This could be because these OTUs facilitate each other thereby making it more difficult for later-colonizing OTUs that specialize in decay to establish. [Alternatively, a high frequency of positively-correlated taxa may indicate that these samples are dominated by co-occurring taxa that are in combat with one another...]
**No pattern apparent**

Hyp: A higher frequency of negatively-correlated OTUs (after controlling for wood trait environmental influences) in a sample community will lead to better-fitting decay models (ne.r2), faster decay (k), and less lagginess (alpha). This could be because the occurance of negatively-correlated OTUs in the same sample indicates that there are negative biotic interactions among members of the initial community, thereby making it easier for later-colonizing OTUs that specialize in decay to establish.
**No pattern apparent**

```{r}

#identify the negative and positive residually-correlated OTU pairs in a given community
#this takes forever.... just read in the derived-data file unless input data changes
#IDcorrelatedOTUsinEachSampl(taxaAndFunguild, comm.otu)
pairsPresent.df<-read.csv(file="derived_data/pairsPresent.csv", row.names=1)


#merge the pairsPresent dataframes with the decay params data and plot
posList<-Plot_signifCor(sign='pos', 
                        valueCol_vec=c("ne.r2", "k", "alpha"), 
                        pairsPresent.df, spdf, seqSamples)
negList<-Plot_signifCor(sign='neg', 
                        valueCol_vec=c("ne.r2", "k", "alpha"), 
                        pairsPresent.df, spdf, seqSamples)

grid.arrange(posList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             posList[['k']] + guides(color=FALSE, shape=FALSE), 
             posList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             negList[['ne.r2']] + guides(color=FALSE, shape=FALSE), 
             negList[['k']] + guides(color=FALSE, shape=FALSE), 
             negList[['alpha']] + guides(color=FALSE, shape=FALSE),
             
             ncol=3)

```

