---
title: "Initial community and wood distances versus decay parameter distances"
date: "10/26/2017"
output: github_document
---

```{r setup, include=FALSE}
#chunk options
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

#libraries
library(knitr) #r markdown
library(readr) #read/write
library(vegan) #ordination stuff
library(tidyr) #data formatting
library(dplyr) #data formatting
library(ggplot2) #plotting
library(gridExtra) #plotting

#fxns
source("code/distance_fxns.R")

#load sample info
stemSamples<-read.csv("derived_data/stemSamples.csv")

#load microbial community
comm.otu<-read.csv("derived_data/comm_otu.csv", row.names=1, stringsAsFactors = FALSE)

seqSamples<-read.csv("derived_data/seqSamples.csv", stringsAsFactors = FALSE)

#load wood trait data
traits.mean<-read.csv("derived_data/traits_code.csv", stringsAsFactors = FALSE)

#load decay trajectories
spdf <- read_csv("derived_data/decayfits.csv")

```

### Plot microbial community distances vs decay param distances between species+size

Community composition, bray
```{r}
pList<-Make_commDistvDist_Fig(distType="bray", valueCol_vec=c("ne.r2","k"), seqSamples, stemSamples, comm.otu, spdf)
grid.arrange(pList[['ne.r2']], pList[['k']])
```

Community composition, jaccard
```{r}
pList<-Make_commDistvDist_Fig(distType="jaccard", valueCol_vec=c("ne.r2","k"), seqSamples, stemSamples, comm.otu, spdf)
grid.arrange(pList[['ne.r2']], pList[['k']])
```

Richness
```{r}
pList<-Make_commDistvDist_Fig(distType="richness", valueCol_vec=c("ne.r2","k"), seqSamples, stemSamples, comm.otu, spdf)
grid.arrange(pList[['ne.r2']], pList[['k']])
```

Not a lot of information in initial mean microbial composition distance (bray or jaccard) about differences in species+size decay trajectories (communities are all really dissimilar). Likewise, richness is not useful.



### Plot wood trait distances vs decay param distances between species+size
```{r}
pList<-Make_woodTraitDistvDist_Fig(valueCol_vec=c("ne.r2","k"), seqSamples, stemSamples, traits.mean, spdf)
grid.arrange(pList[['ne.r2']], pList[['k']])
```

Seems like there is more information in inital wood trait distance about species+size decay trajectories than there is in the initial community composition/richness

